### 2017-06-18 paper.tex
Project created

\documentclass[a4paper]{IEEEconf}

\usepackage{natbib}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}

\usepackage{tikz}
\usetikzlibrary{math,fpu,calc,fit,mindmap,backgrounds,positioning}

\usepackage{xspace}
\newcommand{\eg}{e.g.\xspace}
\newcommand{\etal}{et al.\xspace}
\newcommand{\ie}{i.e.\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\vs}{\textit{vs.}\xspace}

\graphicspath{{figs/}}

\title{PInSoRo: A Dataset for Deep-learning of Human-Robot Social Interactions}

\author{Séverin Lemaignan, Tony Belpaeme}

\date{\today}

\begin{document}
\maketitle

\begin{abstract}

%% Original abstract -- kept for future reference
%Over the last decades, efforts around the world have sought to improve
%human-robot interactions, more recently with a focus on longer-term
%interactions. If worthy, these efforts are often incremental and many
%problems remain. In the light of the recent and rapid progress of machine
%learning, we propose here to shift the paradigm: instead of more and more
%complex symbolic cognitive modeling of the interactions between humans and
%robots, we argue that natural and sustainable human-robot interaction could
%instead be directly learned from actual human-human interactions. This paper
%outlines the main challenges ahead, and proposes a first concrete step
%toward the operationalization of this novel paradigm: the creation of a
%large dataset of human-robot social interactions, suitable for machine
%learning.

Child-robot interactions are increasingly being explored in domains which
require longer-term application, such as healthcare and education. In order
for a robot to behave in an appropriate manner over longer timescales, its
behaviours should be coterminous with that of the interacting children.
Generating such sustained and engaging social behaviours is an on-going research
challenge, and we argue here that the recent progress of deep machine learning
opens new perspectives that the HRI community should embrace. As an initial
step in that direction, we present a new large and open dataset of
children social interactions, designed with deep-learning applications in mind.
Our acquisition methodology relies on little-constrainted \emph{free play}:
by doing so, we seek to capture a rich set of behavioural patterns occurring
in natural social interactions.

\todo[inline]{the dataset is not specific to robotics...PinSoRo is maybe not the
    best name?}

\end{abstract}

\section{Machine Learning: the Next Horizon for Social Robots?}

While the family of \emph{recurrent neural networks} have repeatedly made the
headlines over the last few years with impressive results, notably in image
classification, image labelling and automatic translation, they have been 
largely ignored in many other fields so far as they are perceived to require very
large datasets (hundreds of thousands to millions of observations) to actually
build up useful capabilities. Even though neural networks have demonstrated
compelling results in open-ended, under-defined tasks like image labelling, they
did not stand out as attractive approaches to problems involving high dimensions
with relatively small datasets available -- like human-robot social
interactions.

Besides, if one considers ``social interactions'' to also entail joint
behavioural dynamics, and therefore, some sort of temporal modeling, neural
networks look even less enticing as time is notably absent from most of the
tasks which neural networks have been successful at.

In 2015, the Google DeepMind team demonstrated how a convolutional
recurrent neural network could learn to play the game Break-Out (amongst
48 other Atari games) by only \emph{looking} at the gaming console
screen~\cite{mnih2015human}. This result represents a major milestone: they show
that a relatively small sample size (about 500 games) is sufficient for an 
artificial agent to not only learn how to play (which requires an implicit model 
of time to adequately move the Break-Out paddle), but to also create gaming 
strategies that \emph{look like} they would necessitate planning (the system first
breaks bricks on one side to eventually get the ball to break-out and reach the area
\emph{above} the remaining bricks, therefore ensuring rapid progress in the
game). We argue that the complexity of mechanisms that such a neural network has been able
to quickly uncover and model should invite our
community to question its applicability to human-robot interactions (HRI) in general,
and sustained, natural child-robot interactions in particular.

However, the lack of a widespread HRI dataset suitable for the training of
neural networks is a critical obstacle to this initial exploration. We present
hereafter such a dataset, and detail our acquisition methodology.

\section{Machine learning and social behaviour}

Using interaction datasets to teach robots how to socially behave has been
previously explored, and can be considered as an extension of the traditional
learning from demonstration (LfD) paradigms to social interactions (for
instance~\cite{nehaniv2007imitation,mohammad2015interaction}). Previous examples
have generally focused on low-level recognition or generation of short,
self-standing behaviours, including social gestures~\cite{nagai2005learning} and
gazing behaviours~\cite{calinon2006teaching}.

Based on a human-human interaction dataset, Liu \etal~\cite{liu2014how} have
investigated machine learning approaches to learn longer interaction sequences.
Using unsupervised learning, they train a robot to act as a shop-keeper,
generating both speech and socially acceptable motions. Their approach remains
task-specific, and while they report only limited success, they emphasise the
``life-likeness'' of the generated behaviours.

Kim \etal~\cite{kim2015pororobot} highlight that applying deep learning to
visual scene information in an HRI scenario was successful, but that
generating behaviours for the robot to be able to act in a dynamic and uncertain
environment remains a challenge.
\todo[inline]{cite Ogata's work}

%They rely on unsupervised learning, with a
%training stage consisting in clustering behavioural states (a state consist in
%\emph{behavioural elements} ( joint states of the two interacting agents,
%associated to the current speech elements, as well as the current joint
%\emph{spatial formation}), followed by the {WIP]

These examples show the burgeoning interest of our community for the automatic
learning of social interactions, but also highlight the lack of structure of
these research efforts, as further illustrated by the quasi-absence of public
and large datasets of human-robot interactions. To our best knowledge, only the
$H^3R$ Explanation Corpus~\cite{mohammad2008h} and the Vernissage
Corpus~\cite{jayagopi2013vernissage} have been published to date. The $H^3R$
Explanation Corpus is a human-human and human-robot dataset focusing on a
``assembly/disassembly explanation'' task and includes physiological signals (22
human-robot interactions), but is not publicly available. the Vernissage Corpus
includes one museum guide robot interacting with two people (13 interactions in
total), with recordings and annotations of poses and speech audio (stated to be
publicly available). Both these corpora are however too small for
machine-learning applications.

\todo[inline]{look at child-child datasets!}

\section{The Plymouth Interacting Social Robots Dataset (PInSoRo)}

As introduced, our aim is to build a dataset that can be used to train
machine-learning algorithms to both recognise and generate complex cognitive and
social dynamics in human-robot interactions. The Plymouth Interacting Social
Robots (PInSoRo) Dataset is such a novel dataset, comprising of \emph{natural
social} interactions arising between children, and recorded in a way suitable
for research methodologies based on machine learning. As detailed in
section~\ref{availablity}, this dataset is open and made freely available to any
interested researcher.

\subsection{The freeplay sandbox}

Critically, the acquisition of such a dataset requires first the definition of a
task to be performed by the participants while they are recorded. The nature of
this task influences in fundamental way the kind of interactions that might be
observed, and thus, learned.  Because they usually target very specific social
or cognitive skills in a tightly controlled manner, traditional socio-cognitive tasks are
either toy scenarios (\ie they do not adequately mirror real-world situations)
or simple, constrained tasks that do not reflect the complexity and dynamics of
real-world interactions. For this dataset, we aim on the contrary at a rich set
of naturally-occuring social interactions, and as such we have designed a novel
task that is:

\begin{itemize}
    \item be \textit{fundamentally social}, \ie the task would make little sense for an
        agent alone;
    \item foster rich \textit{multi-modal interaction}: simultaneous speech, gesture, and gaze
        behaviours are to be observed;
    \item exhibit \textit{non-trivial dynamics}, such as implicit turn-taking;
    \item should cover a \textit{broad range of interaction situations}.
\end{itemize}

This new experimental task is based on \emph{free play} interactions. Pairs of
children (4-7 years old) are invited to freely draw and interact with items
displayed on an interactive table, telling and acting stories with no explicit
goal set in advance (Fig.~\ref{fig|setup}).

As a result, the interaction space is high-dimensional, along two main axis: the
game-related actions (drawing, manipulating items) and the social interactions
(looking at the other's actions, skink
ocial interaction situation based on
free dramatic play: 
Critically we propose to limit these tasks to
\emph{face-to-face} social interactions, either dyadic or triadic. This constrains the dataset to
a more tractable domain, and should ensure technical feasibility.

, free play social
interactions exhibit a rich set of natural cognitive and social dynamics that
are desirable.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{setup.png}
    \caption{The free play social interactions sandbox\todo[inline]{replace with a
    cihld-child configuration}}
    \label{fig|setup}
\end{figure}

To mitigate the technical challenges presented by free play interactions in
their most unconstrained version (like
high dimensionality and uncertainties), we have developped a \emph{free play
sandbox}, Figure~\ref{fig|setup}: a large touchscreen (60cm $\times$ 33cm) is
used as an interactive surface (\emph{sandtray}
paradigm~\cite{baxter2012touchscreen}). Two agents (either a pair of children or
one child and one robot) play together by freely moving on a 2D plane items
(characters and blocks, see Figure~\ref{fig|sandbox}). A background image
depicts a generic empty environment, with different symbolic zones (water,
grass, path, bushes...). The players do not have any particular task to
complete, they are simply invited to freely play.

children can engage in open-ended and
non-directive play situations, yet sufficiently
defined to be reproducible;

\subsection{Socio-cognitive Framing}

focus on abstract socio-cognitive facets (robot
perception and manipulation are simplified); well suited for qualitative and
quantitative analysis using metrics like Słowinski’s Individual Motor Signature
(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
interactions, with-me-ness (for assessment of co-engagment).

carefully framed: as far as possible, they ne-
cessitate only the cognitive capabilities that they ev-
idence (e.g. if they evidence a purely socio-linguistic
mechanism, they will not mandate complex action ca-
pabilities – they may benefit from it, though),

This point is especially important as it allows an
incremental path: a robot should be able to tackle one or
several of these task independently of the others, and a re-
search team may progressively extend their cognitive models
to incorporate more and more of the cognitive capabilities
required by the robot to address all of the tasks.

\subsection{Analysis of the free play sandbox}

Our free play sandbox elicits a loosely structured play situation. The
situation is indeed essentially aimless (free play). The goals and play episodes
that we observe during the interactions are not pre-defined, and essentially
unknown until they are decided by the players. In this sense, the free play
sandbox permits and elicits what is fundamentally \emph{unstructured play}.

However, a second look reveals some important structuring elements.

First, the physical bounds of the sandbox (an interactive table) limit the
play zone to a well defined and relatively small area. As a consequence,
children are mostly static (they are sitting in front of the table) and their
primary form of interaction is 2D pick and place of items by drag and dropping
them on the screen.

Second, the game items themselves structure the game scenarios. Items are either
iconic characters (animals or children) or plain blocks
(Figure~\ref{fig|sandbox}). The former have strong semantics associated to them
(like 'crocodiles like water and eat children'), while the later have little
semantics associated. The game background, with its recognizable zones, also
elicit a particular type of games (like building a zoo or pretending we explore
the savannah).

Finally, the social setting (two players facing each other) implicitely invites
social interaction: onlooker play, for instance, is unlikely to be observed as
the children are physically engaged with the game and the other participant
(they are sitting in front of game, close to each other). The three higher forms
of social play listed by Parten (parallel play, associative play and cooperative
play) are however possible with this setup.

Two more elements (the role and place of the experimenter, and the initial
prompt given to the participant) have an important impact on the shape of the
play situation. We discuss separately these two points in the description of the
experimental procedure.

Overall, the free play sandbox supports a \emph{loosely structured} form of play: the
actual play situations are not known and might change several times during the
interaction; the game actions, even though all based on a single modality (picking and
placing game items), are unlimited; the social interactions between participants
are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
unconstrainted. However, the broad domain of the play situations and the range of
possible social interactions is bound by the physical bounds of the play zone
and the theme of the game items.

\subsection{Forms of play elicited by the sandbox}

\begin{itemize}
    \item dramatic play
    \item storytelling
    \item creative play
    \item pretend play
\end{itemize}

\paragraph{From unstructured to structured play}

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{sandbox}
    \caption{Example of a game situation: the child and the robot `build'
    together a zoo. The animals and the blocks can be dragged over the whole
    play area, while the background picture is static and purely decorative.}
    \label{fig|sandbox}
\end{figure}

\paragraph{Pretence}

\begin{figure}
    \centering
    \includegraphics[width=0.7\linewidth]{sandbox-release-cheetah}
    \caption{Possible play situation: the children pretend to release the cheetah
    and create a story where the cheetah scares away the visitors.}
    \label{}
\end{figure}

\subsection{Implementation}

The sandbox is implemented using mainly two frameworks: the Qt's \emph{QtQuick} framework
for the graphical interface of the game, and the \emph{Robot Operating System}
(ROS) for the modular implementation of the data processing and behaviour
generation pipelines.

Figure~\ref{fig|architecture} presents the software architecture of the sandbox.

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{freeplay-sandbox-architecture}
\label{fig|architecture}
\caption{Architecture of the free play sandbox. Yellow nodes represent
    hardware (cameras and robot); purple nodes relate to the game interface;
    red nodes manage the robot; the green node implements the game policy (when
    the robot is playing); blue nodes relates to visual attention tracking.}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{rviz-sandtray}
    \caption{The freeplay sandbox, viewed at runtime within ROS RViz. The
    current background drawing (bottom left) is published on a regular image
    topic. Simple computer vision is used to segment it into zones (visible in
    the central panel). The poses and bounding boxes of the interactive items
    are published as well, and turned into an occupancy map, used to plan the
    robot's arm motion.
    visualisation tool.}
    \label{fig|rviz}
\end{figure}

\section{Dataset Structure}

\subsection{Acquisition Procedure}

\subsection{Initial Dataset Analysis}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{analysis}
    \caption{Screenshot of the interaction analysis tool: faces of the two
    children are recorded, as well as any interaction occuring on the
    interactive table.}
    \label{fig|analysis}
\end{figure}

\subsection{Dataset availablity}
\label{availablity}

The dataset is freely available to any interested researcher. Due to ethical
and data protection regulations, the dataset is however made available in two
forms: a public, Creative-Commons licensed, version that do not include any
material enabling the identification of the children (in particular, it does not
contain any video stream); and a complete version that includes all video
streams. This second version is freely available as well, but interested
researchers must first fill a data protection form.

\section{Research Questions and Hypotheses}

The mechanisms underlying the social interactions happening during free play
raise important questions for social robotics.

\subsection{Game Dynamics}

\begin{itemize}
    \item do we observe a sequence of sub-games?
    \item how are these sub-games initiated?
    \item how does A keep track of what B is playing at?
    \item how do we segment the action flow into meaningful play situations?
\end{itemize}

\subsection{Social Dynamics}

\begin{itemize}
    \item does a ``social protocol'' establish?
    \item if so, what are these social rules?
    \item implicit vs explicit rule setting?
    \item what interaction modalities (or combination thereof) are relied upon
        to establish and maintain this ``social protocol''?
\end{itemize}

\subsection{Mental representations}

\begin{itemize}
    \item Does A know what B is thinking of at time t?
    \item Do the participants perform explicit grounding of their respective mental models
        (ie, ask the other one what she is doing/thinking/planning to do?)
    \item How does the implicit grounding take place? attention tracking?
    \item Do we need at all to know what the other intends to do to actually
        collaborate? surface alignment vs deep grounding
    \item Procedural (how do you do it?) vs Semantic
        grounding (what are you doing? playing at?)
\end{itemize}

\subsection{Interplay between these questions}

\begin{figure}
\centering

\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[
                font=\sffamily,
                >=latex,
                every edge/.style={draw, very thick},
                node/.style={draw, rounded corners, align=center, inner sep=5pt, fill=black!20},
                label/.style={midway, align=center, font=\scriptsize\sffamily, fill=white}]

    %%% NODES
    \node [node] (mentalrepresentations) {Mental Representations};
    \node [above left=of mentalrepresentations] (gameactions) {\it game actions};
    \node [node, above=of gameactions] (gamedynamics) {Game dynamics};
    \node [above right=of mentalrepresentations] (socialactions) {\it social actions};
    \node [node, above=of socialactions ] (socialdynamics) {Social Dynamics};

    %%% CONNECTIONS
    \path (gamedynamics) edge [<->, dashed] node[label, above] {?}(socialdynamics);
    \path (gamedynamics) edge [<-, bend right] node[label] {mediate}(gameactions);
    \path (socialactions) edge [->, bend left] node[label] {reflect}(mentalrepresentations);
    \path (socialactions) edge [->, dashed] node[label,above] {?}(gamedynamics);
    \path (gameactions) edge [->, dashed] node[label,above] {?}(socialdynamics);
    \path (gameactions) edge [->, bend right] node[label] {reflect}(mentalrepresentations);
    \path (socialdynamics) edge [<-, bend left] node[label] {shape}(socialactions);

\end{tikzpicture}
}
\label{fig|researchquestions}
\caption{Game dynamics, social dynamics and mental representations are linked
    with each other through social actions (implicit and explicit communication)
    and actions on the environment (mainly game actions)}
\end{figure}

\subsection{Child-robot baseline}

During the acquisition campaign for the PinSoRo dataset, we have also recorded a
\emph{child-robot} interaction baseline. We were using the exact same protocol
as described above, only with one child replaced by a Nao robot.

The robot was fully autonomous, and programmed to exhibit a non-social behaviour
(its action policy was simply to repeatedly move interactive items to
pre-defined zones, entierly ignoring its child partner).

As such, this behaviour (and the resulting reaction of the children to a
robot acting in a non-social way) represent an experimental baseline against which
future work can be compared.

\section{Conclusion}

By analyzing them from the
perspective of artifical socio-cognition and human-robot interaction, we
evidence key social patterns and show that robots can recognise them,
interpret them, and (where applicable) generate them in a task-agnostic
fashion. By subsequently reusing these social patterns in similar
child-robot free play situations, we show a significant increase of the
interaction duration.

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}

### 2017-06-21 paper.tex
Update on Overleaf.
CoRL template, updated pictures, started restructuring

@@
-\documentclass[a4paper]{IEEEconf}
+\documentclass{article}
+
+\usepackage{corl_2017}
 
@@
 
-\author{Séverin Lemaignan, Tony Belpaeme}
+\author{Séverin Lemaignan, Charlotte Edmunds, Tony Belpaeme}
 
@@
 
-\section{Machine learning and social behaviour}
+\subsection{Machine learning and social behaviour}
 
@@
 
+analysis of behavioural alignment between partners (via
+metrics like the recently proposed \emph{Individual Motor
+Signature}~\cite{slowinski2016dynamic})
+
 This new experimental task is based on \emph{free play} interactions. Pairs of
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{setup.png}
-    \caption{The free play social interactions sandbox\todo[inline]{replace with a
-    cihld-child configuration}}
+    \includegraphics[width=0.4\linewidth]{setup-child-child.png}
+    \hspace{1em}
+    \includegraphics[width=0.4\linewidth]{setup-child-robot.png}
+    \caption{The free play social interactions sandbox. Data acquisition is
+    performed under two (between subjets) conditions: either children playing together, or one
+    child playing with a \emph{non-social} robot.}
     \label{fig|setup}
@@
 
-\subsection{Analysis of the free play sandbox}
-
-Our free play sandbox elicits a loosely structured play situation. The
-situation is indeed essentially aimless (free play). The goals and play episodes
-that we observe during the interactions are not pre-defined, and essentially
-unknown until they are decided by the players. In this sense, the free play
-sandbox permits and elicits what is fundamentally \emph{unstructured play}.
-
-However, a second look reveals some important structuring elements.
-
-First, the physical bounds of the sandbox (an interactive table) limit the
-play zone to a well defined and relatively small area. As a consequence,
-children are mostly static (they are sitting in front of the table) and their
-primary form of interaction is 2D pick and place of items by drag and dropping
-them on the screen.
-
-Second, the game items themselves structure the game scenarios. Items are either
-iconic characters (animals or children) or plain blocks
-(Figure~\ref{fig|sandbox}). The former have strong semantics associated to them
-(like 'crocodiles like water and eat children'), while the later have little
-semantics associated. The game background, with its recognizable zones, also
-elicit a particular type of games (like building a zoo or pretending we explore
-the savannah).
-
-Finally, the social setting (two players facing each other) implicitely invites
-social interaction: onlooker play, for instance, is unlikely to be observed as
-the children are physically engaged with the game and the other participant
-(they are sitting in front of game, close to each other). The three higher forms
-of social play listed by Parten (parallel play, associative play and cooperative
-play) are however possible with this setup.
-
-Two more elements (the role and place of the experimenter, and the initial
-prompt given to the participant) have an important impact on the shape of the
-play situation. We discuss separately these two points in the description of the
-experimental procedure.
-
-Overall, the free play sandbox supports a \emph{loosely structured} form of play: the
-actual play situations are not known and might change several times during the
-interaction; the game actions, even though all based on a single modality (picking and
-placing game items), are unlimited; the social interactions between participants
-are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
-unconstrainted. However, the broad domain of the play situations and the range of
-possible social interactions is bound by the physical bounds of the play zone
-and the theme of the game items.
-
-\subsection{Forms of play elicited by the sandbox}
-
-\begin{itemize}
-    \item dramatic play
-    \item storytelling
-    \item creative play
-    \item pretend play
-\end{itemize}
-
-\paragraph{From unstructured to structured play}
-
 \begin{figure}
@@
     \caption{Example of a game situation: the child and the robot `build'
-    together a zoo. The animals and the blocks can be dragged over the whole
-    play area, while the background picture is static and purely decorative.}
+    together a zoo. Items (animals, characters...) can be dragged over the whole
+    play area, while the background picture can be painted over.}
     \label{fig|sandbox}
-\end{figure}
-
-\paragraph{Pretence}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.7\linewidth]{sandbox-release-cheetah}
-    \caption{Possible play situation: the children pretend to release the cheetah
-    and create a story where the cheetah scares away the visitors.}
-    \label{}
 \end{figure}
@@
 
-Figure~\ref{fig|architecture} presents the software architecture of the sandbox.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{rviz-sandtray}
+    \caption{The freeplay sandbox, viewed at runtime within ROS RViz. The
+    current background drawing is published on a regular image
+    topic and computer vision is used to segment it into zones (visible in
+    the central panel). The poses and bounding boxes of the interactive items
+    are published as well, and turned into an occupancy map, used to plan the
+    robot's arm motion.}
+    \label{fig|rviz}
+\end{figure}
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{freeplay-sandbox-architecture}
-\label{fig|architecture}
-\caption{Architecture of the free play sandbox. Yellow nodes represent
-    hardware (cameras and robot); purple nodes relate to the game interface;
-    red nodes manage the robot; the green node implements the game policy (when
-    the robot is playing); blue nodes relates to visual attention tracking.}
-\end{figure*}
+\section{Dataset Structure}
 
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{rviz-sandtray}
+    \includegraphics[width=0.9\linewidth]{3d-point-cloud-facial-features}
     \caption{The freeplay sandbox, viewed at runtime within ROS RViz. The
@@
 
-\section{Dataset Structure}
+\subsection{Conditions}\label{conditions}
+
+\begin{itemize}
+\item
+  Condition A: one pair of children freely play together on the
+  sandtray.
+\item
+  Condition B: one child play together with the robot, the robot does
+  not exhibit any social behaviour.
+\end{itemize}
+
+\subsection{Demographics}\label{demographics}
+
+\begin{itemize}
+\item
+  Condition A: 40 pairs of children (ie, 80 children)
+\item
+  Condition B: 40 children
+\end{itemize}
+
+2 age groups, balanced across conditions: 4 y.o. (nursery) and 6 to 7
+y.o. (Y1/Y2)
+
+Normally developping children, from local primary schools and nurseries.
+
+\subsection{Experimental Setup}\label{experimental-setup}
+
+On large touchscreen, mounted on a horizontal frame (\emph{sandtray});
+both children (condition A, Figure 2) or the child and the robot
+(condition B, Figure 3) are facing each other.
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.4\linewidth]{setup_child_child_top}
+    \hspace{1.5cm}
+    \includegraphics[width=0.4\linewidth]{setup_child_robot_top}
+    \caption{Experimental setup, child-child condition (left) and child-robot
+    condition (right).}
+\end{figure}
+
+The experimenter stays in the room, visible from the child. When
+requested by the nursery or school, another childminder (nursery/school
+staff) can be present in the room. She/he is however asked not to step
+in during the experiment.
+
+The interaction is recorded by 3 cameras (on `environment' camera
+filming the whole scene; two RGB-D cameras, focused on each of the
+children faces).
+
+Audio is recorded as well with 2 microphones integrated to the RGB-D
+cameras.
+
+One computer (integrated with the touchscreen) manages the game and the
+synchronous recording of the cameras and microphones.
 
@@
 
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  welcome (5 min):
+
+  \begin{itemize}
+  \item
+    place children on cushions
+  \item
+    give them the yellow/purple sport bibs
+  \item
+    make sure hairs are tied (to help with face tracking)
+  \item
+    ask names (but do not write them down)
+  \item
+    fill up questionaire on the tablet -- age, gender, familiarity with
+    tablet
+  \item
+    \textbf{remind the children that they can withdraw at anytime}
+  \item
+    \textbf{explain the purpose of the study: showing robots how
+    children play}
+  \item
+    in condition B, present briefly the robot -- saying it is not a very
+    sociable robot, and it might play alone.
+  \end{itemize}
+\item
+  visual focus task (30 seconds)
+\item
+  tutorial (1-2 min): explain how to interact with the game, ensure the
+  children are confident with the manipulation/drawing
+\item
+  main freeplay task (up to 40 min):
+
+  \begin{itemize}
+  \item
+    start the recording
+  \item
+    example prompt: ``I let you play now! You can build a zoo, or tell
+    stories, as you want''
+  \item
+    let children play
+  \item
+    encourage children to keep playing at least 5 min
+  \item
+    once they wish to stop, stop recording
+  \end{itemize}
+\item
+  debriefing (5 min):
+
+  \begin{itemize}
+  \item
+    answer possible questions from the children
+  \item
+    give stickers
+  \end{itemize}
+\end{enumerate}
+
+Total duration: 30 minutes per group.
 \subsection{Initial Dataset Analysis}
@@
 
+\section{Discussion}
+
+\subsection{Analysis of the free play sandbox}
+
+Our free play sandbox elicits a loosely structured play situation. The
+situation is indeed essentially aimless (free play). The goals and play episodes
+that we observe during the interactions are not pre-defined, and essentially
+unknown until they are decided by the players. In this sense, the free play
+sandbox permits and elicits what is fundamentally \emph{unstructured play}.
+
+However, a second look reveals some important structuring elements.
+
+First, the physical bounds of the sandbox (an interactive table) limit the
+play zone to a well defined and relatively small area. As a consequence,
+children are mostly static (they are sitting in front of the table) and their
+primary form of interaction is 2D pick and place of items by drag and dropping
+them on the screen.
+
+Second, the game items themselves structure the game scenarios. Items are either
+iconic characters (animals or children) or plain blocks
+(Figure~\ref{fig|sandbox}). The former have strong semantics associated to them
+(like 'crocodiles like water and eat children'), while the later have little
+semantics associated. The game background, with its recognizable zones, also
+elicit a particular type of games (like building a zoo or pretending we explore
+the savannah).
+
+Finally, the social setting (two players facing each other) implicitely invites
+social interaction: onlooker play, for instance, is unlikely to be observed as
+the children are physically engaged with the game and the other participant
+(they are sitting in front of game, close to each other). The three higher forms
+of social play listed by Parten (parallel play, associative play and cooperative
+play) are however possible with this setup.
+
+Two more elements (the role and place of the experimenter, and the initial
+prompt given to the participant) have an important impact on the shape of the
+play situation. We discuss separately these two points in the description of the
+experimental procedure.
+
+Overall, the free play sandbox supports a \emph{loosely structured} form of play: the
+actual play situations are not known and might change several times during the
+interaction; the game actions, even though all based on a single modality (picking and
+placing game items), are unlimited; the social interactions between participants
+are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
+unconstrainted. However, the broad domain of the play situations and the range of
+possible social interactions is bound by the physical bounds of the play zone
+and the theme of the game items.
+
+\subsection{Forms of play elicited by the sandbox}
+
+\begin{itemize}
+    \item dramatic play
+    \item storytelling
+    \item creative play
+    \item pretend play
+\end{itemize}
+
 \section{Conclusion}
@@
 
-\bibliographystyle{plain}
 \bibliography{biblio}

### 2017-06-21 paper.tex
WIP on the abstract

@@
 challenge, and we argue here that the recent progress of deep machine learning
-opens new perspectives that the HRI community should embrace. As an initial
-step in that direction, we present a new large and open dataset of
-children social interactions, designed with deep-learning applications in mind.
-Our acquisition methodology relies on little-constrainted \emph{free play}:
-by doing so, we seek to capture a rich set of behavioural patterns occurring
-in natural social interactions.
+opens new perspectives that the Human-Robot Interaction (HRI) community
+should embrace.
 
-\todo[inline]{the dataset is not specific to robotics...PinSoRo is maybe not the
-    best name?}
+As an initial step in that direction, we introduce a novel
+open dataset of children social interactions, designed with
+deep-learning applications in mind. Our acquisition methodology relies on
+engaging and purposefully underspecified \emph{free play} situations. By doing
+so, we capture a rich set of behavioural patterns occurring in natural
+social interactions. Learning to recognise (and generate) automatically such
+behaviours is one of the ambitious foreseen outcome that this dataset enables.
 
@@
 \section{The Plymouth Interacting Social Robots Dataset (PInSoRo)}
+
+\todo[inline]{the dataset is not specific to robotics...PinSoRo is maybe not the
+    best name?}
 

### 2017-06-21 paper.tex
Intro: cite Ogata's work

@@
 
-While the family of \emph{recurrent neural networks} have repeatedly made the
+While the family of recurrent neural networks (RNN) have repeatedly made the
 headlines over the last few years with impressive results, notably in image
-classification, image labelling and automatic translation, they have been 
-largely ignored in many other fields so far as they are perceived to require very
+classification, image labelling and automatic translation, they have been so far 
+largely ignored by  many other fields as they are perceived to require very
 large datasets (hundreds of thousands to millions of observations) to actually
 build up useful capabilities. Even though neural networks have demonstrated
-compelling results in open-ended, under-defined tasks like image labelling, they
+compelling results in open-ended, under-specified tasks like image labelling, they
 did not stand out as attractive approaches to problems involving high dimensions
@@
 \emph{above} the remaining bricks, therefore ensuring rapid progress in the
-game). We argue that the complexity of mechanisms that such a neural network has been able
-to quickly uncover and model should invite our
-community to question its applicability to human-robot interactions (HRI) in general,
-and sustained, natural child-robot interactions in particular.
+game).
 
-However, the lack of a widespread HRI dataset suitable for the training of
-neural networks is a critical obstacle to this initial exploration. We present
-hereafter such a dataset, and detail our acquisition methodology.
+More recently, Ogata's team~\cite{yang2017repeatable} has demonstrated how an
+adequately configured RNN is able to learn a complex robotic task (folding soft
+objects like towels using a dual-arm mobile manipulator) from only \emph{35}
+demonstrations, one demonstration being a $\approx$ 70 seconds-long teleoperated
+sequence. The network inputs are the raw video stream from the head camera and the
+12 DoF of the two arms and the total training dataset comprises of about 28000
+steps of data.
+
+We argue that the complexity of mechanisms that such neural networks have been
+able to uncover and model should invite our community to question its
+applicability to human-robot interactions in general, and sustained, natural
+child-robot interactions in particular.
+
+However, the current lack of an appropriate HRI dataset suitable for the training of
+neural networks hinders this investigation. We present hereafter such a dataset,
+and detail our acquisition methodology.
 
@@
 environment remains a challenge.
-\todo[inline]{cite Ogata's work}
 

### 2017-06-21 paper.tex
WIP

@@
 12 DoF of the two arms and the total training dataset comprises of about 28000
-steps of data.
+steps of data. Successfully folding towels entails an explicit sequencing of
+actions (therefore implicit temporal modeling). The fact that such a complex
+process can be successfully learned from a very small dataset certainly opens
+many prospects.
 
-We argue that the complexity of mechanisms that such neural networks have been
-able to uncover and model should invite our community to question its
+Indeed, we argue that the complexity of mechanisms that such neural networks
+have been able to uncover and model should invite our community to question its
 applicability to human-robot interactions in general, and sustained, natural
 child-robot interactions in particular.
-
 However, the current lack of an appropriate HRI dataset suitable for the training of
@@
 
-\subsection{Machine learning and social behaviour}
+\subsection*{Machine Learning and Social Behaviours}
 
-Using interaction datasets to teach robots how to socially behave has been
-previously explored, and can be considered as an extension of the traditional
+The use of interaction datasets to teach robots how to socially behave has been
+previously explored, and can be considered as the extension of the traditional
 learning from demonstration (LfD) paradigms to social interactions (for
@@
 environment remains a challenge.
-
-%They rely on unsupervised learning, with a
-%training stage consisting in clustering behavioural states (a state consist in
-%\emph{behavioural elements} ( joint states of the two interacting agents,
-%associated to the current speech elements, as well as the current joint
-%\emph{spatial formation}), followed by the {WIP]
+They rely on unsupervised learning, with a
+training stage consisting in clustering behavioural states (a state consist in
+\emph{behavioural elements} (joint states of the two interacting agents,
+associated to the current speech elements, as well as the current joint
+\emph{spatial formation}), followed by the \todo[inline]{finish that}.
 
@@
 learning of social interactions, but also highlight the lack of structure of
-these research efforts, as further illustrated by the quasi-absence of public
-and large datasets of human-robot interactions. To our best knowledge, only the
-$H^3R$ Explanation Corpus~\cite{mohammad2008h} and the Vernissage
-Corpus~\cite{jayagopi2013vernissage} have been published to date. The $H^3R$
-Explanation Corpus is a human-human and human-robot dataset focusing on a
-``assembly/disassembly explanation'' task and includes physiological signals (22
-human-robot interactions), but is not publicly available. the Vernissage Corpus
-includes one museum guide robot interacting with two people (13 interactions in
-total), with recordings and annotations of poses and speech audio (stated to be
-publicly available). Both these corpora are however too small for
-machine-learning applications.
+these research efforts, as further illustrated by the lack of large, open
+datasets of social interactions, suitable for machine-learning applications.
+
+One of the most interesting datasets is the \emph{Multimodal Dyadic Behavior
+Dataset} (\emph{MMDB},~\cite{rehg2013decoding}). It comprises of 160 sessions of
+3 to 5 minute child-adult interactions. During these interactions, the
+experimenter plays with toddlers (1.5 to 2.5 years old) in semi-structured ways
+which was designed to
+elicit a broad range of social behaviors
 

### 2017-06-21 paper.tex
Paragraph on the ethical clearance

@@
 
+All data has been collected by researchers at the Plymouth University, under a
+protocol approved by the university ethics committee. The parents of the
+participants explicitly consented to sharing of their child's video and audio
+with the research community. The data is labelled with a unique participant code
+only and does not contain any identifying information, except the participant's
+images. The child's age and gender are also available.
+
 \begin{enumerate}

### 2017-06-21 paper.tex
Mention the MMDB dataset

@@
 
-One of the most interesting datasets is the \emph{Multimodal Dyadic Behavior
+One interesting dataset is the \emph{Multimodal Dyadic Behavior
 Dataset} (\emph{MMDB},~\cite{rehg2013decoding}). It comprises of 160 sessions of
 3 to 5 minute child-adult interactions. During these interactions, the
-experimenter plays with toddlers (1.5 to 2.5 years old) in semi-structured ways
-which was designed to
-elicit a broad range of social behaviors
+experimenter plays with toddlers (1.5 to 2.5 years old) in semi-structured ways.
+The dataset includes video streams of the faces and the room, audio, physiological data
+(electrodermal activity) as well as manual annotations of certain behaviours
+(like gaze to the examiner, laughter, pointing). The lack of calibration and camera positioning
+information in the dataset limits the possibilities of automatically extracting and
+labelling key interaction features (like facial expressions or joint gaze). Besides,
+this dataset focuses on very young children with in adult-driven interactions,
+and as such does not include any episodes of naturally-occurring social
+interactions between peers.
 
-\todo[inline]{look at child-child datasets!}
+\todo[inline]{look at other child-child datasets!}
+\todo[inline]{mention research on deep-learning to diagnostic autism}
 

### 2017-06-21 paper.tex
Restructuring, WIP on content

@@
 
-As introduced, our aim is to build a dataset that can be used to train
-machine-learning algorithms to both recognise and generate complex cognitive and
-social dynamics in human-robot interactions. The Plymouth Interacting Social
-Robots (PInSoRo) Dataset is such a novel dataset, comprising of \emph{natural
-social} interactions arising between children, and recorded in a way suitable
-for research methodologies based on machine learning. As detailed in
-section~\ref{availablity}, this dataset is open and made freely available to any
-interested researcher.
+The Plymouth Interacting Social Robots (PInSoRo) Dataset comprises of natural
+social interactions arising between children, and recorded in a way suitable
+for research methodologies based on machine learning.
 
-\subsection{The freeplay sandbox}
+As detailed in section~\ref{availability}, this dataset is open and made freely
+available to any interested researcher. It is intended to be used to train
+machine-learning algorithms (typically, deep recurrent network) to both
+recognise and generate complex cognitive and social dynamics occurring in the
+wild during social interactions.
 
-Critically, the acquisition of such a dataset requires first the definition of a
+\subsection{The Freeplay Sandbox}
+
+The acquisition of such a dataset requires first the definition of a
 task to be performed by the participants while they are recorded. The nature of
@@
 real-world interactions. For this dataset, we aim on the contrary at a rich set
-of naturally-occuring social interactions, and as such we have designed a novel
-task that is:
+of naturally-occurring social interactions, and as such we have designed a novel
+task that:
 
 \begin{itemize}
-    \item be \textit{fundamentally social}, \ie the task would make little sense for an
+    \item is fundamentally social, \ie the task would make little sense for an
         agent alone;
-    \item foster rich \textit{multi-modal interaction}: simultaneous speech, gesture, and gaze
+    \item elicit a large range of interaction situations;
+    \item foster rich multi-modal interaction: simultaneous speech, gesture, and gaze
         behaviours are to be observed;
-    \item exhibit \textit{non-trivial dynamics}, such as implicit turn-taking;
-    \item should cover a \textit{broad range of interaction situations}.
+    \item foster non-trivial social dynamics, such as implicit turn-taking.
 \end{itemize}
-
-analysis of behavioural alignment between partners (via
-metrics like the recently proposed \emph{Individual Motor
-Signature}~\cite{slowinski2016dynamic})
 
@@
 
-\subsection{Socio-cognitive Framing}
+\subsubsection{Socio-cognitive Framing}
 
@@
 
-\subsection{Implementation}
+\subsubsection{Implementation}
 
@@
 
-\section{Dataset Structure}
+\subsection{Dataset Acquisition}
+
+All data has been collected by researchers at the Plymouth University, under a
+protocol approved by the university ethics committee. The parents of the
+participants explicitly consented to sharing of their child's video and audio
+with the research community. The data is labelled with a unique participant code
+only and does not contain any identifying information, except the participant's
+images. The child's age and gender are also available.
 
@@
 
-\subsection{Conditions}\label{conditions}
+\subsubsection{Conditions}\label{conditions}
 
@@
 
-\subsection{Demographics}\label{demographics}
+\subsubsection{Demographics}\label{demographics}
 
@@
 
-\subsection{Experimental Setup}\label{experimental-setup}
+\subsubsection{Experimental Setup}\label{experimental-setup}
 
@@
 
-\subsection{Acquisition Procedure}
-
-All data has been collected by researchers at the Plymouth University, under a
-protocol approved by the university ethics committee. The parents of the
-participants explicitly consented to sharing of their child's video and audio
-with the research community. The data is labelled with a unique participant code
-only and does not contain any identifying information, except the participant's
-images. The child's age and gender are also available.
+\subsubsection{Acquisition Procedure}
 
@@
 Total duration: 30 minutes per group.
-\subsection{Initial Dataset Analysis}
+
+\subsubsection{Initial Dataset Analysis}
+
+analysis of behavioural alignment between partners (via
+metrics like the recently proposed \emph{Individual Motor
+Signature}~\cite{slowinski2016dynamic})
 
@@
 
-\subsection{Dataset availablity}
-\label{availablity}
+\section{Dataset Structure}
+
+\subsection{Dataset availability}
+\label{availability}
 

### 2017-06-21 paper.tex
More restructuring

@@
 
-\section{Research Questions and Hypotheses}
+\section{Discussion}
 
-The mechanisms underlying the social interactions happening during free play
-raise important questions for social robotics.
-
-\subsection{Game Dynamics}
-
-\begin{itemize}
-    \item do we observe a sequence of sub-games?
-    \item how are these sub-games initiated?
-    \item how does A keep track of what B is playing at?
-    \item how do we segment the action flow into meaningful play situations?
-\end{itemize}
-
-\subsection{Social Dynamics}
-
-\begin{itemize}
-    \item does a ``social protocol'' establish?
-    \item if so, what are these social rules?
-    \item implicit vs explicit rule setting?
-    \item what interaction modalities (or combination thereof) are relied upon
-        to establish and maintain this ``social protocol''?
-\end{itemize}
-
-\subsection{Mental representations}
-
-\begin{itemize}
-    \item Does A know what B is thinking of at time t?
-    \item Do the participants perform explicit grounding of their respective mental models
-        (ie, ask the other one what she is doing/thinking/planning to do?)
-    \item How does the implicit grounding take place? attention tracking?
-    \item Do we need at all to know what the other intends to do to actually
-        collaborate? surface alignment vs deep grounding
-    \item Procedural (how do you do it?) vs Semantic
-        grounding (what are you doing? playing at?)
-\end{itemize}
-
-\subsection{Interplay between these questions}
+\subsection{Specific research questions}
 
@@
 
-\resizebox{\linewidth}{!}{%
+\resizebox{0.8\linewidth}{!}{%
 \begin{tikzpicture}[
@@
 
+The mechanisms underlying the social interactions happening during free play
+raise important questions for social robotics.
+
+\subsubsection{Game Dynamics}
+
+\begin{itemize}
+    \item do we observe a sequence of sub-games?
+    \item how are these sub-games initiated?
+    \item how does A keep track of what B is playing at?
+    \item how do we segment the action flow into meaningful play situations?
+\end{itemize}
+
+\subsubsection{Social Dynamics}
+
+\begin{itemize}
+    \item does a ``social protocol'' establish?
+    \item if so, what are these social rules?
+    \item implicit vs explicit rule setting?
+    \item what interaction modalities (or combination thereof) are relied upon
+        to establish and maintain this ``social protocol''?
+\end{itemize}
+
+\subsubsection{Mental representations}
+
+\begin{itemize}
+    \item Does A know what B is thinking of at time t?
+    \item Do the participants perform explicit grounding of their respective mental models
+        (ie, ask the other one what she is doing/thinking/planning to do?)
+    \item How does the implicit grounding take place? attention tracking?
+    \item Do we need at all to know what the other intends to do to actually
+        collaborate? surface alignment vs deep grounding
+    \item Procedural (how do you do it?) vs Semantic
+        grounding (what are you doing? playing at?)
+\end{itemize}
+
 \subsection{Child-robot baseline}
 
-During the acquisition campaign for the PinSoRo dataset, we have also recorded a
+During the acquisition campaign for the PInSoRo dataset, we have also recorded a
 \emph{child-robot} interaction baseline. We were using the exact same protocol
@@
 (its action policy was simply to repeatedly move interactive items to
-pre-defined zones, entierly ignoring its child partner).
+pre-defined zones, entirely ignoring its child partner).
 
@@
 future work can be compared.
-
-\section{Discussion}
 

### 2017-06-23 paper.tex
Update on Overleaf.
Added the list of recorded streams

@@
 \usepackage[colorinlistoftodos]{todonotes}
+\usepackage{booktabs}
 
@@
 
+\todo[inline]{mention robot greetings: expectation that the robot can talk
+(and likely, hear)}
+\todo[inline]{mention idea box}
+\todo[inline]{mention robot calibration phase (between 0 and }
+
 Total duration: 30 minutes per group.
@@
 
-\section{Dataset Structure}
+\section{The PInSoRo Dataset}
+
+\subsection{Structure}
+
+The dataset consists in a collection of records. Each record correspond to one
+play interaction between two children. To date (June 2017) \todo[inline]{put
+latest nb} records have been acquired. At the end of the acquisition campaign
+(July 2017), the dataset is planned to include 50 records (\ie 100 children). As
+mentioned above, the duration of each play episode varies ($M=17m13s,
+SD=8m02s$), but it is capped to a maximum of 40 minutes.
+
+\begin{table}[]
+\centering
+\caption{List of datastreams stored in each record}
+\label{table|datastreams}
+\begin{tabular}{@{}lll@{}}
+\toprule
+\bf Domain  & \bf Type                              & \bf Details                              \\ \midrule
+child 1     & audio                                 & 16kHz, mono, semi-directional        \\
+            & face (RGB)                            & qHD (960$\times$540), 30Hz           \\
+            & face (depth)                          & VGA (640$\times$480), 30Hz           \\
+            & facial features                       & 68 3D points (point cloud)           \\ \midrule
+child 2     & audio                                 & 16kHz, mono, semi-directional        \\
+            & face (RGB)                            & qHD (960$\times$540), 30Hz           \\
+            & face (depth)                          & VGA (640$\times$480), 30Hz           \\
+            & facial features                       & 68 3D points (point cloud)           \\ \midrule
+environment & RGB                                   & qHD (960$\times$540), 29.7Hz         \\ \midrule
+touchscreen & background drawing (RGB)              & 4Hz                                  \\
+            & touches                               & 6 points multi-touch                 \\
+            & items position and orientation        &                                      \\ \midrule
+others      & \multicolumn{2}{l}{annotations of social behaviours and remarkable events}   \\
+            & \multicolumn{2}{l}{static transforms between touchscreen and facial cameras} \\
+            & \multicolumn{2}{l}{cameras calibration informations}                         \\ \bottomrule
+\end{tabular}
+\end{table}
 

### 2017-06-23 paper.tex
More on the dataset structure

@@
     \includegraphics[width=0.9\linewidth]{3d-point-cloud-facial-features}
-    \caption{The freeplay sandbox, viewed at runtime within ROS RViz. The
-    current background drawing (bottom left) is published on a regular image
-    topic. Simple computer vision is used to segment it into zones (visible in
-    the central panel). The poses and bounding boxes of the interactive items
-    are published as well, and turned into an occupancy map, used to plan the
-    robot's arm motion.
-    visualisation tool.}
-    \label{fig|rviz}
+    \caption{Reconstructed 3D point cloud with 68 facial features high-lighted}
+    \label{fig|pointcloud}
 \end{figure}
@@
 The dataset consists in a collection of records. Each record correspond to one
-play interaction between two children. To date (June 2017) \todo[inline]{put
-latest nb} records have been acquired. At the end of the acquisition campaign
+play interaction between two children. To date (June 2017) 19 records have been acquired. At the end of the acquisition campaign
 (July 2017), the dataset is planned to include 50 records (\ie 100 children). As
@@
 SD=8m02s$), but it is capped to a maximum of 40 minutes.
+
+\todo[inline]{update with last nb before submitting}
+
+Data is collected using the ROS middleware\footnote{\url{http://ros.org}} and
+recorded as \emph{bag} files. Every dataframe is timestamped.
+Table~\ref{table|datastreams} lists all the recorded datastreams.
 
@@
 \end{table}
+
+As the data is recorded using ROS's bag files, it can be replayed in the exact
+same conditions as it was during the recording. All the video streams use
+calibrated cameras. While the raw RGB and depth streams are neither rectified
+nor registered together (due to performance considerations during the
+recording), it can be easily achieved as a post-process step, using standard ROS
+tools (the ROS {\tt rgbd\_pipeline}). Pre-configured scripts are available
+alongside the
+dataset\footnote{Available online from
+\url{https://github.com/severin-lemaignan/freeplay-sandbox-analysis}.} to
+republish registered streams and the corresponding 3D point-clouds (as seen in
+Figure~\ref{fig|pointcloud}).
 

### 2017-06-26 paper.tex
WIP

@@
 \todo[inline]{look at other child-child datasets!}
-\todo[inline]{mention research on deep-learning to diagnostic autism}
+\todo[inline]{mention research on deep-learning to diagnose autism}
 
@@
 The Plymouth Interacting Social Robots (PInSoRo) Dataset comprises of natural
-social interactions arising between children, and recorded in a way suitable
-for research methodologies based on machine learning.
+social interactions arising between children, and recorded with machine learning
+applications in mind.
 
@@
 available to any interested researcher. It is intended to be used to train
-machine-learning algorithms (typically, deep recurrent network) to both
+machine-learning algorithms (typically, deep recurrent networks) to both
 recognise and generate complex cognitive and social dynamics occurring in the
@@
 
-The acquisition of such a dataset requires first the definition of a
-task to be performed by the participants while they are recorded. The nature of
-this task influences in fundamental way the kind of interactions that might be
+\begin{figure}
+    \centering
+    \includegraphics[width=0.4\linewidth]{setup-child-child.png}
+    \hspace{1em}
+    \includegraphics[width=0.4\linewidth]{child-child-env}
+    \caption{The free play social interactions sandbox: two children interact in
+    Da free-play situation, by drawing and manipulating items on a touchscreen.}
+    \label{fig|setup}
+\end{figure}
+
+The acquisition of such a dataset requires first the definition of a suitable
+task for the participants to perform while they are recorded. The nature of
+this task influences in fundamental ways the kind of interactions that might be
 observed, and thus, learned. Because they usually target very specific social
 or cognitive skills in a tightly controlled manner, traditional socio-cognitive tasks are
-either toy scenarios (\ie they do not adequately mirror real-world situations)
-or simple, constrained tasks that do not reflect the complexity and dynamics of
-real-world interactions. For this dataset, we aim on the contrary at a rich set
-of naturally-occurring social interactions, and as such we have designed a novel
-task that:
+usually simple, constrained tasks that do not reflect the complexity and dynamics of
+real-world interactions. On the contrary, this dataset aims at capturing a rich set
+of naturally-occurring social interactions. As such, a suitable task has to be:
 
 \begin{itemize}
-    \item is fundamentally social, \ie the task would make little sense for an
-        agent alone;
     \item elicit a large range of interaction situations;
@@
         behaviours are to be observed;
+    \item fundamentally social, \ie the task would make little sense for an
+        agent alone;
     \item foster non-trivial social dynamics, such as implicit turn-taking.
@@
 
-This new experimental task is based on \emph{free play} interactions. Pairs of
+We present a new experimental task based on \emph{free play} interactions. Pairs of
 children (4-7 years old) are invited to freely draw and interact with items
@@
 
-As a result, the interaction space is high-dimensional, along two main axis: the
-game-related actions (drawing, manipulating items) and the social interactions
-(looking at the other's actions, skink
-ocial interaction situation based on
-free dramatic play: 
-Critically we propose to limit these tasks to
-\emph{face-to-face} social interactions, either dyadic or triadic. This constrains the dataset to
-a more tractable domain, and should ensure technical feasibility.
+In order to constraint the dataset to a tractable domain, the task is limited to
+a dyadic \emph{face-to-face} interaction.  The resulting interaction space is
+nevertheless high-dimensional, with two main axis: the game-related actions
+(drawing, manipulating items, telling stories) and the social interactions
+(monitoring the other's actions, discussing and agreeing on joint actions,
+helping behaviours, etc.).
 
-, free play social
-interactions exhibit a rich set of natural cognitive and social dynamics that
-are desirable.
+The \emph{freeplay sandbox} follows the sandtray
+paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
+multitouch) is used as an interactive surface (\emph{sandtray}). The two children play together
+by freely moving interactive items (Fig.~\ref{fig|sandbox}) on the surface. A background image
+depicts a generic empty environment, with different symbolic zones (water,
+grass, path, bushes...). The players do not have any particular task to
+complete, they are simply invited to freely play.
+is implemented wit To mitigate the technical challenges presented by free play interactions in
+their most unconstrained version (like
+high dimensionality and uncertainties), we have developped a \emph{free play
+sandbox}, Fig.~\ref{fig|setup}:
 
@@
     \centering
-    \includegraphics[width=0.4\linewidth]{setup-child-child.png}
-    \hspace{1em}
-    \includegraphics[width=0.4\linewidth]{setup-child-robot.png}
-    \caption{The free play social interactions sandbox. Data acquisition is
-    performed under two (between subjets) conditions: either children playing together, or one
-    child playing with a \emph{non-social} robot.}
-    \label{fig|setup}
+    \includegraphics[width=0.9\linewidth]{sandbox}
+    \caption{Example of a game situation: the two children `build'
+    together a zoo. Items (animals, characters...) can be dragged over the whole
+    play area, while the background picture can be painted over by picking a
+    colour.}
+    \label{fig|sandbox}
 \end{figure}
-
-To mitigate the technical challenges presented by free play interactions in
-their most unconstrained version (like
-high dimensionality and uncertainties), we have developped a \emph{free play
-sandbox}, Figure~\ref{fig|setup}: a large touchscreen (60cm $\times$ 33cm) is
-used as an interactive surface (\emph{sandtray}
-paradigm~\cite{baxter2012touchscreen}). Two agents (either a pair of children or
-one child and one robot) play together by freely moving on a 2D plane items
-(characters and blocks, see Figure~\ref{fig|sandbox}). A background image
-depicts a generic empty environment, with different symbolic zones (water,
-grass, path, bushes...). The players do not have any particular task to
-complete, they are simply invited to freely play.
 
@@
 required by the robot to address all of the tasks.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\linewidth]{sandbox}
-    \caption{Example of a game situation: the child and the robot `build'
-    together a zoo. Items (animals, characters...) can be dragged over the whole
-    play area, while the background picture can be painted over.}
-    \label{fig|sandbox}
-\end{figure}
 

### 2017-06-27 paper.tex
Update on Overleaf.

@@
 \item
-  Condition A: one pair of children freely play together on the
+  Condition A: one pair of children play freely  simultaneously on the
   sandtray.
 \item
-  Condition B: one child play together with the robot, the robot does
+  Condition B: one child plays on the sandtray with the robot. The robot does
   not exhibit any social behaviour.
@@
 \item
-  Condition A: 40 pairs of children (ie, 80 children)
+  Condition A: 40 pairs of children (i.e., 80 children)
 \item
@@
 
-Normally developping children, from local primary schools and nurseries.
+Normally developing children, from local primary schools and nurseries.
 
@@
 On large touchscreen, mounted on a horizontal frame (\emph{sandtray});
-both children (condition A, Figure 2) or the child and the robot
-(condition B, Figure 3) are facing each other.
+both children (Condition A, Figure 5 left) or the child and the robot
+(Condition B, Figure 5 right) are facing each other.
 
@@
 
-The experimenter stays in the room, visible from the child. When
+The experimenter stays in the room, visible to the child. When
 requested by the nursery or school, another childminder (nursery/school
@@
 
-The interaction is recorded by 3 cameras (on `environment' camera
-filming the whole scene; two RGB-D cameras, focused on each of the
-children faces).
-
-Audio is recorded as well with 2 microphones integrated to the RGB-D
-cameras.
+The interaction is recorded by 3 cameras: one `environment' camera
+filming the whole scene and two RGB-D cameras, one focused on each of the
+children's faces. Audio is recorded using the microphones integrated to the RGB-D cameras.
 
@@
 \item
-  welcome (5 min):
+  Welcome (5 min):
 
   \begin{itemize}
-  \item
-    place children on cushions
-  \item
-    give them the yellow/purple sport bibs
-  \item
-    make sure hairs are tied (to help with face tracking)
-  \item
-    ask names (but do not write them down)
-  \item
-    fill up questionaire on the tablet -- age, gender, familiarity with
-    tablet
-  \item
-    \textbf{remind the children that they can withdraw at anytime}
   \item
@@
     sociable robot, and it might play alone.
+  \item
+    place children on cushions
+  \item
+    give them the yellow/purple sport bibs
+  \item
+    make sure their hair is tied back (to help with face tracking)
+  \item
+    ask names (but do not write them down)
+  \item
+     complete questionnaire on the tablet -- age, gender, familiarity with
+    tablet
+  \item
+    \textbf{remind the children that they can withdraw at anytime}
+
   \end{itemize}

### 2017-06-27 paper.tex
Update on Overleaf.
CERE checked protocol

@@
     \item
-    in condition B, present briefly the robot -- saying it is not a very
-    sociable robot, and it might play alone.
+    briefly present the robot -- it gives a short message dependent on condition (Condition A: "Hello, I'm Nao.
+Today, I'll be watching you while you play and soon I'll be able to play with you.
+Is that alright?"; Condition B: "Hello, I'm Nao.
+Today I'll be playing with you.
+Exciting!")
   \item
@@
 \item
-  visual focus task (30 seconds)
+  Visual focus task (30 seconds)
 \item
-  tutorial (1-2 min): explain how to interact with the game, ensure the
+  Tutorial (1-2 min): explain how to interact with the game, ensure the
   children are confident with the manipulation/drawing
+%CERE Are we mentioning the object placement task?
 \item
-  main freeplay task (up to 40 min):
+  Freeplay task (up to 40 min):
 
@@
   \item 
-    example prompt: ``I let you play now! You can build a zoo, or tell
-    stories, as you want''
+     Condition B only: Set up and localise robot (1-3 min)
+  \item
+    example prompt: ``Just to remind you, you can use the animals or draw. Whatever you like. 
+    If you run out of ideas, there's also an ideas box. For example, the first one is a zoo. You could draw a zoo or tell a story. 
+    When you get bored or 
+    don't want to play anymore, just let me know."
   \item
@@
 \item
-  debriefing (5 min):
+  Debriefing (5 min):
 
@@
   \item
-    give stickers
+    give stickers as a thank you
   \end{itemize}
@@
 
-Total duration: 30 minutes per group.
+Total duration: between 20-50 min.
 

### 2017-06-27 paper.tex
WIP

@@
     \caption{The free play social interactions sandbox: two children interact in
-    Da free-play situation, by drawing and manipulating items on a touchscreen.}
-    \label{fig|setup}
+    a free-play situation, by drawing and manipulating items on a touchscreen.}
+    \label{fig|freeplay}
 \end{figure}
@@
 
-We present a new experimental task based on \emph{free play} interactions. Pairs of
-children (4-7 years old) are invited to freely draw and interact with items
+We present a new experimental task based on \emph{free play} interactions. Pairs
+of children (4-7 years old) are invited to freely draw and interact with items
 displayed on an interactive table, telling and acting stories with no explicit
-goal set in advance (Fig.~\ref{fig|setup}).
+goal set in advance (Fig.~\ref{fig|freeplay}). The task is designed so that
+children can engage in open-ended and non-directive play situations, yet
+sufficiently well defined to be controlled, suitable for detailled recording,
+and easily reproducible.
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=0.8\linewidth]{sandbox}
+    \caption{Example of a possible game situation. Items (animals,
+    characters...) can be dragged over the whole play area, while the background
+    picture can be painted over by picking a colour.}
+
+    \label{fig|sandbox}
+\end{figure}
+
+The \emph{freeplay sandbox} follows the sandtray
+paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
+with multitouch support) is used as an interactive surface (\emph{sandtray}). The two children play together
+by freely moving interactive items (Fig.~\ref{fig|sandbox}) on the surface. A background image
+depicts a generic empty environment, with different symbolic colours (water,
+grass, beach, bushes...). By drawing on top of the background picture, the
+children can also alter this environment to their liking. The players do not have any particular task to
+complete, they are simply invited to freely play for as long as they wish.
 
@@
 
-The \emph{freeplay sandbox} follows the sandtray
-paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
-multitouch) is used as an interactive surface (\emph{sandtray}). The two children play together
-by freely moving interactive items (Fig.~\ref{fig|sandbox}) on the surface. A background image
-depicts a generic empty environment, with different symbolic zones (water,
-grass, path, bushes...). The players do not have any particular task to
-complete, they are simply invited to freely play.
-is implemented wit To mitigate the technical challenges presented by free play interactions in
-their most unconstrained version (like
-high dimensionality and uncertainties), we have developped a \emph{free play
-sandbox}, Fig.~\ref{fig|setup}:
+%
+%\paragraph{Socio-cognitive framing of the task}
+%
+%
+%focus on abstract socio-cognitive facets (robot
+%perception and manipulation are simplified); well suited for qualitative and
+%quantitative analysis using metrics like Słowinski’s Individual Motor Signature
+%(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
+%interactions, with-me-ness (for assessment of co-engagment).
+%
+%carefully framed: as far as possible, they necessitate only the cognitive
+%capabilities that they evidence (e.g. if they evidence a purely socio-linguistic
+%mechanism, they will not mandate complex action capabilities – they may benefit
+%from it, though),
+%
+%This point is especially important as it allows an incremental path: a robot
+%should be able to tackle one or several of these task independently of the
+%others, and a research team may progressively extend their cognitive models to
+%incorporate more and more of the cognitive capabilities required by the robot to
+%address all of the tasks.
 
-\begin{figure}
+\subsection{Dataset structure}
+
+The dataset consists in a collection of records. Each record correspond to one
+play interaction between two children. To date (June 2017) 19 records have been acquired. At the end of the acquisition campaign
+(July 2017), the dataset is planned to include 50 records (\ie 100 children). As
+mentioned above, the duration of each play episode varies ($M=17m13s,
+SD=8m02s$), but it is capped to a maximum of 40 minutes.
+
+\todo[inline]{update with last nb before submitting}
+
+Data is collected using the ROS middleware\footnote{\url{http://ros.org}} and
+recorded as \emph{bag} files. Every dataframe is timestamped.
+Table~\ref{table|datastreams} lists all the recorded datastreams.
+
+\begin{table}[]
 \centering
-    \includegraphics[width=0.9\linewidth]{sandbox}
-    \caption{Example of a game situation: the two children `build'
-    together a zoo. Items (animals, characters...) can be dragged over the whole
-    play area, while the background picture can be painted over by picking a
-    colour.}
-    \label{fig|sandbox}
+\caption{List of datastreams stored in each record}
+\label{table|datastreams}
+\begin{tabular}{@{}lll@{}}
+\toprule
+\bf Domain  & \bf Type                              & \bf Details                          \\ \midrule
+child 1     & audio                                 & 16kHz, mono, semi-directional        \\
+            & face (RGB)                            & qHD (960$\times$540), 30Hz           \\
+            & face (depth)                          & VGA (640$\times$480), 30Hz           \\
+            & facial features                       & 68 3D points (point cloud)           \\ \midrule
+child 2     & audio                                 & 16kHz, mono, semi-directional        \\
+            & face (RGB)                            & qHD (960$\times$540), 30Hz           \\
+            & face (depth)                          & VGA (640$\times$480), 30Hz           \\
+            & facial features                       & 68 3D points (point cloud)           \\ \midrule
+environment & RGB                                   & qHD (960$\times$540), 29.7Hz         \\ \midrule
+touchscreen & background drawing (RGB)              & 4Hz                                  \\
+            & touches                               & 6 points multi-touch                 \\
+            & items position and orientation        &                                      \\ \midrule
+others      & \multicolumn{2}{l}{annotations of social behaviours and remarkable events}   \\
+            & \multicolumn{2}{l}{static transforms between touchscreen and facial cameras} \\
+            & \multicolumn{2}{l}{cameras calibration informations}                         \\ \bottomrule
+\end{tabular}
+\end{table}
+
+As the data is recorded using ROS's bag files, it can be replayed in the exact
+same conditions as it was during the recording. All the video streams use
+calibrated cameras. Only the raw RGB and depth video streams are recorded. While
+the raw RGB and depth streams are neither rectified nor registered together (due
+to performance considerations during the recording), it can be easily achieved
+as a post-process step, using standard ROS tools (the ROS {\tt rgbd\_pipeline}).
+Pre-configured scripts are available alongside the dataset\footnote{Available
+online from
+\url{https://github.com/severin-lemaignan/freeplay-sandbox-analysis}.} to
+republish registered streams and the corresponding 3D point-clouds (as seen in
+Fig.~\ref{fig|rviz}, left).
+
+\subsection{Hardware and Software}
+
+\paragraph{Hardware}
+The sandtray is made of a 27" Samsung All-In-One computer running Ubuntu Linux
+and equipped with a fast 1TB SSD hard-drive. The computer is held horizontal in
+a custom aluminium frame standing 26cm above the ground
+(Fig.~\ref{fig|freeplay}, right). All the cameras are plugged directly over
+USB-3 to the computer. The computer performs all the data acquisition using ROS
+Kinetic.
+
+The children faces are recorded from two Intel RealSense SR300 RGB-D cameras
+(short range, 0.2m to 1.2m, indoor RGB-D cameras), placed at the corners of the
+the sandtray (Figs.~\ref{fig|rviz} and~\ref{fig|setup}), and tilted to face the
+children. The cameras are rigidly mounted on custom 3D-printed brackets, and
+their 6D pose with respect to the sandtray screen is precisely known and
+recorded (this enables in particular accurate estimation of the children's head
+pose).
+
+Audio is recorded from the same SR300 cameras (one audio stream is recorded for
+each child, from the camera facing her/him).
+
+Finally, a third RGB camera (the RGB stream of a Microsoft Kinect One) records
+the whole interaction setting. These video stream is intended to support human
+coders while annontating the interaction, and is not precisely calibrated.
+
+\begin{figure}[ht!]
+    \centering
+    \includegraphics[width=0.5\linewidth]{rviz-sandtray}
+    \hspace{0.2em}
+    \includegraphics[width=0.48\linewidth]{3d-point-cloud-facial-features}
+    \caption{The freeplay sandbox, visualised at runtime with ROS RViz. The
+    current background drawing is published on a regular image
+    topic and computer vision is used to segment it into zones (visible in
+    the central panel, right picture). The poses of the interactive items
+    are published, as well as an estimate of the children's head pose.
+    The left picture shows a reconstructed 3D point cloud of one child face,
+    along with the detected 68 facial features.}
+    \label{fig|rviz}
 \end{figure}
 
-children can engage in open-ended and
-non-directive play situations, yet sufficiently
-defined to be reproducible;
+\paragraph{Software}
+The sandbox is implemented with two frameworks: the Qt's \emph{QtQuick} framework
+for the graphical interface of the game, and the \emph{Robot Operating System}
+(ROS) for the modular implementation of the data processing and data acquisition
+pipelines. A dedicated bridge between QtQuick and ROS has been specifically
+developped to enable the game interface to export the positions of every
+iteractive items as they move, the background image whenever it is painted over,
+and the children's touches. The game interface is open-source and available online:
+\url{https://github.com/severin-lemaignan/freeplay-sandbox-qt}.
 
-\subsubsection{Socio-cognitive Framing}
+By relying on ROS for the data acquisition, real-time monitoring of the
+interaction is also possible (Fig.~\ref{fig|rviz}).
+The ROS data acquisition pipeline is open-source as well, and available from
+\url{https://github.com/severin-lemaignan/freeplay-sandbox-ros}.
 
-focus on abstract socio-cognitive facets (robot
-perception and manipulation are simplified); well suited for qualitative and
-quantitative analysis using metrics like Słowinski’s Individual Motor Signature
-(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
-interactions, with-me-ness (for assessment of co-engagment).
-
-carefully framed: as far as possible, they ne-
-cessitate only the cognitive capabilities that they ev-
-idence (e.g. if they evidence a purely socio-linguistic
-mechanism, they will not mandate complex action ca-
-pabilities – they may benefit from it, though),
-
-This point is especially important as it allows an
-incremental path: a robot should be able to tackle one or
-several of these task independently of the others, and a re-
-search team may progressively extend their cognitive models
-to incorporate more and more of the cognitive capabilities
-required by the robot to address all of the tasks.
-
-\subsubsection{Implementation}
-
-The sandbox is implemented using mainly two frameworks: the Qt's \emph{QtQuick} framework
-for the graphical interface of the game, and the \emph{Robot Operating System}
-(ROS) for the modular implementation of the data processing and behaviour
-generation pipelines.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\linewidth]{rviz-sandtray}
-    \caption{The freeplay sandbox, viewed at runtime within ROS RViz. The
-    current background drawing is published on a regular image
-    topic and computer vision is used to segment it into zones (visible in
-    the central panel). The poses and bounding boxes of the interactive items
-    are published as well, and turned into an occupancy map, used to plan the
-    robot's arm motion.}
-    \label{fig|rviz}
-\end{figure}
+Finally, we have developped a web-based \emph{supervisor} that permits the
+experimenter to remotely start/stop the ROS nodes and the game GUI, as well as to
+record annotations during the experiment. The supervisor ensures that the exact
+same recording procedure (detailled in the next section) is followed for every
+participants. The supervisor is available online as well:
+\url{https://github.com/severin-lemaignan/freeplay-sandbox-supervisor}.
 
@@
 images. The child's age and gender are also available.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\linewidth]{3d-point-cloud-facial-features}
-    \caption{Reconstructed 3D point cloud with 68 facial features high-lighted}
-    \label{fig|pointcloud}
-\end{figure}
 
@@
     condition (right).}
+    \label{fig|setup}
 \end{figure}
@@
 
-\section{The PInSoRo Dataset}
-
-\subsection{Structure}
-
-The dataset consists in a collection of records. Each record correspond to one
-play interaction between two children. To date (June 2017) 19 records have been acquired. At the end of the acquisition campaign
-(July 2017), the dataset is planned to include 50 records (\ie 100 children). As
-mentioned above, the duration of each play episode varies ($M=17m13s,
-SD=8m02s$), but it is capped to a maximum of 40 minutes.
-
-\todo[inline]{update with last nb before submitting}
-
-Data is collected using the ROS middleware\footnote{\url{http://ros.org}} and
-recorded as \emph{bag} files. Every dataframe is timestamped.
-Table~\ref{table|datastreams} lists all the recorded datastreams.
-
-\begin{table}[]
-\centering
-\caption{List of datastreams stored in each record}
-\label{table|datastreams}
-\begin{tabular}{@{}lll@{}}
-\toprule
-\bf Domain  & \bf Type                              & \bf Details                          \\ \midrule
-child 1     & audio                                 & 16kHz, mono, semi-directional        \\
-            & face (RGB)                            & qHD (960$\times$540), 30Hz           \\
-            & face (depth)                          & VGA (640$\times$480), 30Hz           \\
-            & facial features                       & 68 3D points (point cloud)           \\ \midrule
-child 2     & audio                                 & 16kHz, mono, semi-directional        \\
-            & face (RGB)                            & qHD (960$\times$540), 30Hz           \\
-            & face (depth)                          & VGA (640$\times$480), 30Hz           \\
-            & facial features                       & 68 3D points (point cloud)           \\ \midrule
-environment & RGB                                   & qHD (960$\times$540), 29.7Hz         \\ \midrule
-touchscreen & background drawing (RGB)              & 4Hz                                  \\
-            & touches                               & 6 points multi-touch                 \\
-            & items position and orientation        &                                      \\ \midrule
-others      & \multicolumn{2}{l}{annotations of social behaviours and remarkable events}   \\
-            & \multicolumn{2}{l}{static transforms between touchscreen and facial cameras} \\
-            & \multicolumn{2}{l}{cameras calibration informations}                         \\ \bottomrule
-\end{tabular}
-\end{table}
-
-As the data is recorded using ROS's bag files, it can be replayed in the exact
-same conditions as it was during the recording. All the video streams use
-calibrated cameras. While the raw RGB and depth streams are neither rectified
-nor registered together (due to performance considerations during the
-recording), it can be easily achieved as a post-process step, using standard ROS
-tools (the ROS {\tt rgbd\_pipeline}). Pre-configured scripts are available
-alongside the
-dataset\footnote{Available online from
-\url{https://github.com/severin-lemaignan/freeplay-sandbox-analysis}.} to
-republish registered streams and the corresponding 3D point-clouds (as seen in
-Figure~\ref{fig|pointcloud}).
-
 \subsection{Dataset availability}
@@
 
-\section{Discussion}
+\section{Applications}
 

### 2017-06-27 paper.tex
WIP section on annotations of social behaviours

@@
 
-\subsubsection{Initial Dataset Analysis}
+\subsection{Annotations of Social Behaviours}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\linewidth]{analysis}
+    \caption{Screenshot of the interaction analysis tool: key record information
+    are played back in a synchronised (faces and audio of the two children,
+    interaction occuring on the interactive table, etc.) while the human coder
+    annotate the record timeline with social situations of interest.}
+    \label{fig|analysis}
+\end{figure}
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{analysis}
-    \caption{Screenshot of the interaction analysis tool: faces of the two
-    children are recorded, as well as any interaction occuring on the
-    interactive table.}
-    \label{fig|analysis}
-\end{figure}
+Słowinski’s Individual Motor Signature
+(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
+interactions,
+
+\cite{olswang2006reliability}
 
@@
 \subsection{Specific research questions}
-
-\begin{figure}
-\centering
-
-\resizebox{0.8\linewidth}{!}{%
-\begin{tikzpicture}[
-                font=\sffamily,
-                >=latex,
-                every edge/.style={draw, very thick},
-                node/.style={draw, rounded corners, align=center, inner sep=5pt, fill=black!20},
-                label/.style={midway, align=center, font=\scriptsize\sffamily, fill=white}]
-
-    %%% NODES
-    \node [node] (mentalrepresentations) {Mental Representations};
-    \node [above left=of mentalrepresentations] (gameactions) {\it game actions};
-    \node [node, above=of gameactions] (gamedynamics) {Game dynamics};
-    \node [above right=of mentalrepresentations] (socialactions) {\it social actions};
-    \node [node, above=of socialactions ] (socialdynamics) {Social Dynamics};
-
-    %%% CONNECTIONS
-    \path (gamedynamics) edge [<->, dashed] node[label, above] {?}(socialdynamics);
-    \path (gamedynamics) edge [<-, bend right] node[label] {mediate}(gameactions);
-    \path (socialactions) edge [->, bend left] node[label] {reflect}(mentalrepresentations);
-    \path (socialactions) edge [->, dashed] node[label,above] {?}(gamedynamics);
-    \path (gameactions) edge [->, dashed] node[label,above] {?}(socialdynamics);
-    \path (gameactions) edge [->, bend right] node[label] {reflect}(mentalrepresentations);
-    \path (socialdynamics) edge [<-, bend left] node[label] {shape}(socialactions);
-
-\end{tikzpicture}
-}
-\label{fig|researchquestions}
-\caption{Game dynamics, social dynamics and mental representations are linked
-    with each other through social actions (implicit and explicit communication)
-    and actions on the environment (mainly game actions)}
-\end{figure}
-
-The mechanisms underlying the social interactions happening during free play
-raise important questions for social robotics.
+%
+%\begin{figure}
+%\centering
+%
+%\resizebox{0.8\linewidth}{!}{%
+%\begin{tikzpicture}[
+%                font=\sffamily,
+%                >=latex,
+%                every edge/.style={draw, very thick},
+%                node/.style={draw, rounded corners, align=center, inner sep=5pt, fill=black!20},
+%                label/.style={midway, align=center, font=\scriptsize\sffamily, fill=white}]
+%
+%    %%% NODES
+%    \node [node] (mentalrepresentations) {Mental Representations};
+%    \node [above left=of mentalrepresentations] (gameactions) {\it game actions};
+%    \node [node, above=of gameactions] (gamedynamics) {Game dynamics};
+%    \node [above right=of mentalrepresentations] (socialactions) {\it social actions};
+%    \node [node, above=of socialactions ] (socialdynamics) {Social Dynamics};
+%
+%    %%% CONNECTIONS
+%    \path (gamedynamics) edge [<->, dashed] node[label, above] {?}(socialdynamics);
+%    \path (gamedynamics) edge [<-, bend right] node[label] {mediate}(gameactions);
+%    \path (socialactions) edge [->, bend left] node[label] {reflect}(mentalrepresentations);
+%    \path (socialactions) edge [->, dashed] node[label,above] {?}(gamedynamics);
+%    \path (gameactions) edge [->, dashed] node[label,above] {?}(socialdynamics);
+%    \path (gameactions) edge [->, bend right] node[label] {reflect}(mentalrepresentations);
+%    \path (socialdynamics) edge [<-, bend left] node[label] {shape}(socialactions);
+%
+%\end{tikzpicture}
+%}
+%\label{fig|researchquestions}
+%\caption{Game dynamics, social dynamics and mental representations are linked
+%    with each other through social actions (implicit and explicit communication)
+%    and actions on the environment (mainly game actions)}
+%\end{figure}
+%
+%The mechanisms underlying the social interactions happening during free play
+%raise important questions for social robotics.
 

### 2017-06-27 paper.tex
Section on data post processing

@@
 same conditions as it was during the recording. All the video streams use
-calibrated cameras. Only the raw RGB and depth video streams are recorded. While
-the raw RGB and depth streams are neither rectified nor registered together (due
-to performance considerations during the recording), it can be easily achieved
-as a post-process step, using standard ROS tools (the ROS {\tt rgbd\_pipeline}).
-Pre-configured scripts are available alongside the dataset\footnote{Available
-online from
-\url{https://github.com/severin-lemaignan/freeplay-sandbox-analysis}.} to
-republish registered streams and the corresponding 3D point-clouds (as seen in
-Fig.~\ref{fig|rviz}, left).
+calibrated cameras. Only the raw RGB and depth video streams are recorded.
 
@@
 
-\subsubsection{Conditions}\label{conditions}
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\linewidth]{setup_child_child_top}
+    %\hspace{1.5cm}
+    %\includegraphics[width=0.4\linewidth]{setup_child_robot_top}
+    %\caption{Experimental setup, child-child condition (left) and child-robot condition (right).}
+    \caption{Experimental setup}
+    \label{fig|setup}
+\end{figure}
 
-\begin{itemize}
-\item
-  Condition A: one pair of children play freely  simultaneously on the
-  sandtray.
-\item
-  Condition B: one child plays on the sandtray with the robot. The robot does
-  not exhibit any social behaviour.
-\end{itemize}
-
-\subsubsection{Demographics}\label{demographics}
-
-\begin{itemize}
-\item
-  Condition A: 40 pairs of children (i.e., 80 children)
-\item
-  Condition B: 40 children
-\end{itemize}
+The experimenter stays in the room, visible to the child. When
+requested by the nursery or school, another childminder (nursery/school
+staff) can be present in the room. She/he is however asked not to step
+in during the experiment.
 
@@
 
-\begin{figure}[htbp]
-    \centering
-    \includegraphics[width=0.4\linewidth]{setup_child_child_top}
-    \hspace{1.5cm}
-    \includegraphics[width=0.4\linewidth]{setup_child_robot_top}
-    \caption{Experimental setup, child-child condition (left) and child-robot
-    condition (right).}
-    \label{fig|setup}
-\end{figure}
-
-The experimenter stays in the room, visible to the child. When
-requested by the nursery or school, another childminder (nursery/school
-staff) can be present in the room. She/he is however asked not to step
-in during the experiment.
-
-The interaction is recorded by 3 cameras: one `environment' camera
-filming the whole scene and two RGB-D cameras, one focused on each of the
-children's faces. Audio is recorded using the microphones integrated to the RGB-D cameras.
-
-One computer (integrated with the touchscreen) manages the game and the
-synchronous recording of the cameras and microphones.
-
 \subsubsection{Acquisition Procedure}
@@
 \begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
 \item
@@
 Total duration: between 20-50 min.
+
+\subsection{Data post-processing}
+
+While
+the raw RGB and depth streams are neither rectified nor registered together (due
+to performance considerations during the recording), it can be easily achieved
+as a post-process step, using standard ROS tools (the ROS {\tt rgbd\_pipeline}).
+Pre-configured scripts are available alongside the dataset\footnote{Available
+online from
+\url{https://github.com/severin-lemaignan/freeplay-sandbox-analysis}.} to
+republish registered streams and the corresponding 3D point-clouds (as seen in
+Fig.~\ref{fig|rviz}, left).
+
+Using the {\tt gazr} face tracking library\footnote{Available online from
+\url{https://github.com/severin-lemaignan/gazr}. {\tt gazr} relies on the {\tt
+dlib} face tracker for facial features
+extraction.}~\cite{lemaignan2016realtime}, children faces and facial features can
+are extracted and reprojected onto the 3D point-clouds (Fig.~\ref{fig|rviz},
+right). Faces have been successfully detected in 57\% of the recorded frames
+in the dataset. Undetected faces are due to occlusions (the child put her arm in
+front of the camera), large head rotations or quick child motion.
 

### 2017-06-27 paper.tex
WIP

@@
 applicability to human-robot interactions in general, and sustained, natural
-child-robot interactions in particular.
-However, the current lack of an appropriate HRI dataset suitable for the training of
-neural networks hinders this investigation. We present hereafter such a dataset,
-and detail our acquisition methodology.
+child-robot interactions in particular.  However, the current lack of an
+appropriate HRI dataset suitable for the training of neural networks hinders
+this investigation. We present hereafter such a dataset, and detail our
+acquisition methodology.
 
@@
     \centering
-    \includegraphics[width=0.4\linewidth]{setup-child-child.png}
+    \includegraphics[width=0.55\linewidth]{setup-child-child.png}
     \hspace{1em}
@@
 The dataset consists in a collection of records. Each record correspond to one
-play interaction between two children. To date (June 2017) 19 records have been acquired. At the end of the acquisition campaign
-(July 2017), the dataset is planned to include 50 records (\ie 100 children). As
-mentioned above, the duration of each play episode varies ($M=17m13s,
-SD=8m02s$), but it is capped to a maximum of 40 minutes.
-
-\todo[inline]{update with last nb before submitting}
+play interaction between two children. To date (June 2017) 25 records have been acquired (\ie 50 children). At the end of the acquisition campaign
+(July 2017), the dataset is planned to include 50 records. As
+mentioned above, the duration of each play episode varies ($M=20m51s,
+SD=10m40s$), but it is capped to a maximum of 40 minutes.
 
@@
             & touches                               & 6 points multi-touch                 \\
-            & items position and orientation        &                                      \\ \midrule
+            & items position and orientation        & ($x,y,\theta$)                       \\ \midrule
 others      & \multicolumn{2}{l}{annotations of social behaviours and remarkable events}   \\
@@
 
-All data has been collected by researchers at the Plymouth University, under a
-protocol approved by the university ethics committee. The parents of the
-participants explicitly consented to sharing of their child's video and audio
-with the research community. The data is labelled with a unique participant code
-only and does not contain any identifying information, except the participant's
-images. The child's age and gender are also available.
-
 \begin{figure}[htbp]
@@
     %\caption{Experimental setup, child-child condition (left) and child-robot condition (right).}
-    \caption{Experimental setup}
+    \caption{Experimental setup. Children are facing each other and sit on
+    cushions. Each child wears a bright sport bib, either purple or yellow, to
+    facilitate later identification.}
     \label{fig|setup}
@@
 
-\subsubsection{Experimental Setup}\label{experimental-setup}
+\paragraph{Acquisition Procedure}
 
-On large touchscreen, mounted on a horizontal frame (\emph{sandtray});
-both children (Condition A, Figure 5 left) or the child and the robot
-(Condition B, Figure 5 right) are facing each other.
+The following acquisition procedure has been followed with all participants.
 
-\subsubsection{Acquisition Procedure}
-
-\begin{enumerate}
-\item
-  Welcome (5 min):
-
-  \begin{itemize}
-  \item
-    \textbf{explain the purpose of the study: showing robots how
-    children play}
-    \item
-    briefly present the robot -- it gives a short message dependent on condition (Condition A: "Hello, I'm Nao.
-Today, I'll be watching you while you play and soon I'll be able to play with you.
-Is that alright?"; Condition B: "Hello, I'm Nao.
-Today I'll be playing with you.
-Exciting!")
-  \item
-    place children on cushions
-  \item
-    give them the yellow/purple sport bibs
-  \item
-    make sure their hair is tied back (to help with face tracking)
-  \item
-    ask names (but do not write them down)
-  \item
-     complete questionnaire on the tablet -- age, gender, familiarity with
-    tablet
-  \item
-    \textbf{remind the children that they can withdraw at anytime}
-
-  \end{itemize}
-\item
-  Visual focus task (30 seconds)
-\item
-  Tutorial (1-2 min): explain how to interact with the game, ensure the
-  children are confident with the manipulation/drawing
-%CERE Are we mentioning the object placement task?
-\item
-  Freeplay task (up to 40 min):
-
-  \begin{itemize}
-  \item
-    start the recording
-  \item 
-     Condition B only: Set up and localise robot (1-3 min)
-  \item
-    example prompt: ``Just to remind you, you can use the animals or draw. Whatever you like. 
-    If you run out of ideas, there's also an ideas box. For example, the first one is a zoo. You could draw a zoo or tell a story. 
-    When you get bored or 
-    don't want to play anymore, just let me know."
-  \item
-    let children play
-  \item
-    encourage children to keep playing at least 5 min
-  \item
-    once they wish to stop, stop recording
-  \end{itemize}
-\item
-  Debriefing (5 min):
-
-  \begin{itemize}
-  \item
-    answer possible questions from the children
-  \item
-    give stickers as a thank you
-  \end{itemize}
-\end{enumerate}
-
-\todo[inline]{mention robot greetings: expectation that the robot can talk
-(and likely, hear)}
-\todo[inline]{mention idea box}
-\todo[inline]{mention robot calibration phase (between 0 and }
-
-Total duration: between 20-50 min.
+\newcommand{\tabitem}{~~\llap{\textbullet}~~}
+    \begin{tabular}{@{}p{0.2\linewidth}p{0.8\linewidth}@{}}
+\toprule
+\bf Welcome                       & \tabitem explain the purpose of the study: showing robots how children play  \\
+\emph{(about 5 min)}              & \tabitem briefly present a Nao robot -- the robot stands up, gives a short
+                                    message (\emph{"Hello, I'm Nao. Today, I'll be watching you while you
+                                    play and soon I'll be able to play with you. Is that alright?"}), and
+                                    sit down. The robot is then put aside and does not intervene anymore
+                                    for the remaining of the experiment; \\
+                                    %Condition B: "Hello, I'm Nao. Today I'll be playing with you. Exciting!")\\ 
+                                  & \tabitem place children on cushions  \\ 
+                                  & \tabitem give them the yellow/purple sport
+                                  bibs and make sure their hair is tied \\ 
+                                  & \tabitem complete demographics on the tablet \\
+                                  & \tabitem remind the children that they can withdraw at anytime \\ \midrule
+\bf Visual focus task             &  \\ 
+\emph{(30 sec)}                   &  \\ \midrule
+\bf Tutorial                      & explain how to interact with the game, ensure the children are confident with the manipulation/drawing \\ 
+\emph{(1-2 min)}                  &  \\ \midrule
+\bf Item placement game           &  \\
+\emph{(5 to 10 min)}              &  \\ \midrule
+\bf Freeplay task                 & \tabitem start the recording \\
+\emph{(up to 40 min)}             & \tabitem initial prompt: \emph{"Just to remind you, you can use the animals or draw. Whatever you
+                                  like. If you run out of ideas, there's also an ideas box. For example, the first one is a
+                                  zoo. You could draw a zoo or tell a story. When you get bored or don't want to play
+                                  anymore, just let me know."} \\
+                                  & \tabitem let children play \\
+                                  & \tabitem once they wish to stop, stop recording \\ \midrule
+\bf Debriefing                    &  \tabitem answer possible questions from the children \\
+\emph{(about 2 min)}              & \tabitem give stickers as a thank you \\ \bottomrule
+\end{tabular}
 
@@
 
+The last part of the dataset consist in annotations of the
+children's social behaviours. Some of these behaviours are automatically
+computed and coded (mutual gaze, for instance), some are human-coded.
+
+The annotations are integrated in the dataset as a timestamped datastream.
+
+\paragraph{Human-coded social behaviours}
+
+We use the \emph{Social Communication Coding System} proposed by
+Olswang~\etal\cite{olswang2006reliability}. It consists in 6 constructs
+characterising social communication (table~\ref{table|sccs}).
+
+% Please add the following required packages to your document preamble:
+% \usepackage{booktabs}
+\begin{table}[]
+\centering
+\caption{Constructs of the Social Communication Coding System (SCCS), taken from~\cite{olswang2006reliability}}
+\label{table|sccs}
+\renewcommand{\arraystretch}{1.8}
+\begin{tabular}{@{}lp{0.8\linewidth}@{}}
+\toprule
+\bf Construct          & \bf Description                                                                                                                                                                                                                                                                                                                                           \\ \midrule
+Hostile/coercive   & anti-social behavior ; examples include grabbing, hitting, taunting, provoking, yelling.                                                       \\
+Prosocial/engaged  & engaged in a positive social interaction ; examples include on-task work or conversations, helping, sharing, and compromising.                 \\
+Assertive          & child stating a position or opinion ; examples include asserting beliefs (verbally or nonverbally), directing the partner.                     \\
+Passive/disengaged & lack of involvement in the, activity, physical disengagement ; examples include staring into space, putting head down, speaking quite softly, and walking away from an activity.\\
+Adult seeking      & effort to seek assistance or attention from an adult ; examples include requesting help.                                                       \\
+Irrelevant         & child being actively engaged in an off-task behavior ; examples include starting a different activity, fiddling with objects as a main focus of attention, goofy or silly behaviors.
+\\ \bottomrule
+\end{tabular}
+\end{table}
+
 \begin{figure}
@@
 
+Several metrics and coding scheme
+
 analysis of behavioural alignment between partners (via
@@
 
-\cite{olswang2006reliability}
-
-\subsection{Dataset availability}
+\subsection{Ethical considerations and dataset availability}
 \label{availability}
 
-The dataset is freely available to any interested researcher. Due to ethical
-and data protection regulations, the dataset is however made available in two
-forms: a public, Creative-Commons licensed, version that do not include any
-material enabling the identification of the children (in particular, it does not
-contain any video stream); and a complete version that includes all video
+All data has been collected by researchers at the Plymouth University, under a
+protocol approved by the university ethics committee. The parents of the
+participants explicitly consented to sharing of their child's video and audio
+with the research community. The data is labelled with a unique participant code
+only and does not contain any identifying information, except the participant's
+images. The child's age and gender are also available.
+
+The dataset is freely available to any interested researcher\footnote{Note to
+the reviewers: the link to the dataset's dedicated website will be added to the
+final version of the article.}. Due to ethical and data protection regulations,
+the dataset is however made available in two forms: a public, Creative-Commons
+licensed, version that do not include any video material of the children (no
+video streams, audio included); and a complete version that includes all video
 streams. This second version is freely available as well, but interested
@@
 
-\section{Applications}
-
-\subsection{Specific research questions}
+%\section{Applications}
+%
+%\subsection{Specific research questions}
 %
@@
 %raise important questions for social robotics.
-
-\subsubsection{Game Dynamics}
-
-\begin{itemize}
-    \item do we observe a sequence of sub-games?
-    \item how are these sub-games initiated?
-    \item how does A keep track of what B is playing at?
-    \item how do we segment the action flow into meaningful play situations?
-\end{itemize}
-
-\subsubsection{Social Dynamics}
-
-\begin{itemize}
-    \item does a ``social protocol'' establish?
-    \item if so, what are these social rules?
-    \item implicit vs explicit rule setting?
-    \item what interaction modalities (or combination thereof) are relied upon
-        to establish and maintain this ``social protocol''?
-\end{itemize}
-
-\subsubsection{Mental representations}
-
-\begin{itemize}
-    \item Does A know what B is thinking of at time t?
-    \item Do the participants perform explicit grounding of their respective mental models
-        (ie, ask the other one what she is doing/thinking/planning to do?)
-    \item How does the implicit grounding take place? attention tracking?
-    \item Do we need at all to know what the other intends to do to actually
-        collaborate? surface alignment vs deep grounding
-    \item Procedural (how do you do it?) vs Semantic
-        grounding (what are you doing? playing at?)
-\end{itemize}
+%
+%
+%\subsubsection{Game Dynamics}
+%
+%\begin{itemize}
+%    \item do we observe a sequence of sub-games?
+%    \item how are these sub-games initiated?
+%    \item how does A keep track of what B is playing at?
+%    \item how do we segment the action flow into meaningful play situations?
+%\end{itemize}
+%
+%\subsubsection{Social Dynamics}
+%
+%\begin{itemize}
+%    \item does a ``social protocol'' establish?
+%    \item if so, what are these social rules?
+%    \item implicit vs explicit rule setting?
+%    \item what interaction modalities (or combination thereof) are relied upon
+%        to establish and maintain this ``social protocol''?
+%\end{itemize}
+%
+%\subsubsection{Mental representations}
+%
+%\begin{itemize}
+%    \item Does A know what B is thinking of at time t?
+%    \item Do the participants perform explicit grounding of their respective mental models
+%        (ie, ask the other one what she is doing/thinking/planning to do?)
+%    \item How does the implicit grounding take place? attention tracking?
+%    \item Do we need at all to know what the other intends to do to actually
+%        collaborate? surface alignment vs deep grounding
+%    \item Procedural (how do you do it?) vs Semantic
+%        grounding (what are you doing? playing at?)
+%\end{itemize}
 
@@
 
-\subsection{Analysis of the free play sandbox}
+%\subsection{Analysis of the free play sandbox}
+%
+%Our free play sandbox elicits a loosely structured play situation. The
+%situation is indeed essentially aimless (free play). The goals and play episodes
+%that we observe during the interactions are not pre-defined, and essentially
+%unknown until they are decided by the players. In this sense, the free play
+%sandbox permits and elicits what is fundamentally \emph{unstructured play}.
+%
+%However, a second look reveals some important structuring elements.
+%
+%First, the physical bounds of the sandbox (an interactive table) limit the
+%play zone to a well defined and relatively small area. As a consequence,
+%children are mostly static (they are sitting in front of the table) and their
+%primary form of interaction is 2D pick and place of items by drag and dropping
+%them on the screen.
+%
+%Second, the game items themselves structure the game scenarios. Items are either
+%iconic characters (animals or children) or plain blocks
+%(Figure~\ref{fig|sandbox}). The former have strong semantics associated to them
+%(like 'crocodiles like water and eat children'), while the later have little
+%semantics associated. The game background, with its recognizable zones, also
+%elicit a particular type of games (like building a zoo or pretending we explore
+%the savannah).
+%
+%Two more elements (the role and place of the experimenter, and the initial
+%prompt given to the participant) have an important impact on the shape of the
+%play situation. We discuss separately these two points in the description of the
+%experimental procedure.
+%
+%Overall, the free play sandbox supports a \emph{loosely structured} form of play: the
+%actual play situations are not known and might change several times during the
+%interaction; the game actions, even though all based on a single modality (picking and
+%placing game items), are unlimited; the social interactions between participants
+%are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
+%unconstrainted. However, the broad domain of the play situations and the range of
+%possible social interactions is bound by the physical bounds of the play zone
+%and the theme of the game items.
+%
+%\subsection{Forms of play elicited by the sandbox}
+%
+%\begin{itemize}
+%    \item dramatic play
+%    \item storytelling
+%    \item creative play
+%    \item pretend play
+%\end{itemize}
+%
 
-Our free play sandbox elicits a loosely structured play situation. The
-situation is indeed essentially aimless (free play). The goals and play episodes
-that we observe during the interactions are not pre-defined, and essentially
-unknown until they are decided by the players. In this sense, the free play
-sandbox permits and elicits what is fundamentally \emph{unstructured play}.
+\section{Conclusion}
+%
+%
+%By analyzing them from the
+%perspective of artifical socio-cognition and human-robot interaction, we
+%evidence key social patterns and show that robots can recognise them,
+%interpret them, and (where applicable) generate them in a task-agnostic
+%fashion. By subsequently reusing these social patterns in similar
+%child-robot free play situations, we show a significant increase of the
+%interaction duration.
 
-However, a second look reveals some important structuring elements.
+We introduce a novel dataset for the study of social cognition. The dataset
+consists in dyadic free-play episodes, with children aged 4 to 8.
 
-First, the physical bounds of the sandbox (an interactive table) limit the
-play zone to a well defined and relatively small area. As a consequence,
-children are mostly static (they are sitting in front of the table) and their
-primary form of interaction is 2D pick and place of items by drag and dropping
-them on the screen.
-
-Second, the game items themselves structure the game scenarios. Items are either
-iconic characters (animals or children) or plain blocks
-(Figure~\ref{fig|sandbox}). The former have strong semantics associated to them
-(like 'crocodiles like water and eat children'), while the later have little
-semantics associated. The game background, with its recognizable zones, also
-elicit a particular type of games (like building a zoo or pretending we explore
-the savannah).
+The dataset is build with machine learning applications in mind, and
+particularly, applications to socio-cognitive robotics.
 
@@
 
-Two more elements (the role and place of the experimenter, and the initial
-prompt given to the participant) have an important impact on the shape of the
-play situation. We discuss separately these two points in the description of the
-experimental procedure.
-
-Overall, the free play sandbox supports a \emph{loosely structured} form of play: the
-actual play situations are not known and might change several times during the
-interaction; the game actions, even though all based on a single modality (picking and
-placing game items), are unlimited; the social interactions between participants
-are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
-unconstrainted. However, the broad domain of the play situations and the range of
-possible social interactions is bound by the physical bounds of the play zone
-and the theme of the game items.
-
-\subsection{Forms of play elicited by the sandbox}
-
-\begin{itemize}
-    \item dramatic play
-    \item storytelling
-    \item creative play
-    \item pretend play
-\end{itemize}
-
-\section{Conclusion}
-
-By analyzing them from the
-perspective of artifical socio-cognition and human-robot interaction, we
-evidence key social patterns and show that robots can recognise them,
-interpret them, and (where applicable) generate them in a task-agnostic
-fashion. By subsequently reusing these social patterns in similar
-child-robot free play situations, we show a significant increase of the
-interaction duration.
-
 \bibliography{biblio}

### 2017-06-27 paper.tex
Update on Overleaf.

@@
 %outlines the main challenges ahead, and proposes a first concrete step
-%toward the operationalization of this novel paradigm: the creation of a
+%toward the operationalisation of this novel paradigm: the creation of a
 %large dataset of human-robot social interactions, suitable for machine
@@
 for a robot to behave in an appropriate manner over longer timescales, its
-behaviours should be coterminous with that of the interacting children.
+behaviours should be aligned and meaningful to the interacting children.
 Generating such sustained and engaging social behaviours is an on-going research
@@
 
-As an initial step in that direction, we introduce a novel
+As an initial step in that direction, we introduce a novel,
 open dataset of children social interactions, designed with
-deep-learning applications in mind. Our acquisition methodology relies on
-engaging and purposefully underspecified \emph{free play} situations. By doing
+machine learning applications in mind. Our data acquisition methodology relies on
+a engaging and purposefully underspecified \emph{free play} interaction. By doing
 so, we capture a rich set of behavioural patterns occurring in natural
-social interactions. Learning to recognise (and generate) automatically such
-behaviours is one of the ambitious foreseen outcome that this dataset enables.
+social interactions. Learning to recognise, and possibly generate, such
+behaviours is one of the future research direction that we foresee this dataset to effectively support.
 

### 2017-06-27 paper.tex
Polishing

@@
 recurrent neural network could learn to play the game Break-Out (amongst
-48 other Atari games) by only \emph{looking} at the gaming console
+48 other Atari games) by only looking at the gaming console
 screen~\cite{mnih2015human}. This result represents a major milestone: they show
@@
 of time to adequately move the Break-Out paddle), but to also create gaming 
-strategies that \emph{look like} they would necessitate planning (the system first
-breaks bricks on one side to eventually get the ball to break-out and reach the area
-\emph{above} the remaining bricks, therefore ensuring rapid progress in the
+strategies that look like they would necessitate planning (the system first
+breaks bricks on one side to eventually get the ball to break-out and reach out the area
+above the remaining bricks, therefore ensuring rapid progress in the
 game).
@@
 objects like towels using a dual-arm mobile manipulator) from only \emph{35}
-demonstrations, one demonstration being a $\approx$ 70 seconds-long teleoperated
-sequence. The network inputs are the raw video stream from the head camera and the
-12 DoF of the two arms and the total training dataset comprises of about 28000
-steps of data. Successfully folding towels entails an explicit sequencing of
-actions (therefore implicit temporal modeling). The fact that such a complex
-process can be successfully learned from a very small dataset certainly opens
-many prospects.
+demonstrations of $\approx$ 70 seconds-long teleoperated
+sequences. The network inputs are the raw video stream from the head camera and the
+12 DoF of the two arms. Successfully folding towels entails an explicit sequencing of
+actions (therefore implicit temporal modeling): the fact that such a complex
+process can be successfully learned from a small training dataset should lead
+us to reconsider the range of domains to which RNNs would be applicable.
 
-Indeed, we argue that the complexity of mechanisms that such neural networks
-have been able to uncover and model should invite our community to question its
-applicability to human-robot interactions in general, and sustained, natural
-child-robot interactions in particular.  However, the current lack of an
-appropriate HRI dataset suitable for the training of neural networks hinders
-this investigation. We present hereafter such a dataset, and detail our
-acquisition methodology.
+Indeed, as we have previously argued~\cite{lemaignan2016towards}, we believe
+that the complexity of mechanisms that such neural networks have been able to
+uncover and model should invite our community to question its applicability to
+human-robot interactions in general, and sustained, natural child-robot
+interactions in particular. The current lack of appropriate HRI
+datasets suitable for neural networks training is however an obstacle: we
+introduce in this article the PInSoRo dataset, a large dataset of child-child
+natural interactions suitable for machine learning.
 
@@
 learning from demonstration (LfD) paradigms to social interactions (for
-instance~\cite{nehaniv2007imitation,mohammad2015interaction}). Previous examples
-have generally focused on low-level recognition or generation of short,
-self-standing behaviours, including social gestures~\cite{nagai2005learning} and
-gazing behaviours~\cite{calinon2006teaching}.
+instance~\cite{nehaniv2007imitation,mohammad2015interaction}). Existing
+research focuses however on low-level recognition or generation of short,
+self-standing behaviours, including social gestures~\cite{nagai2005learning}
+and gazing behaviours~\cite{calinon2006teaching}.
 
@@
 ``life-likeness'' of the generated behaviours.
-
-Kim \etal~\cite{kim2015pororobot} highlight that applying deep learning to
-visual scene information in an HRI scenario was successful, but that
-generating behaviours for the robot to be able to act in a dynamic and uncertain
-environment remains a challenge.
-They rely on unsupervised learning, with a
-training stage consisting in clustering behavioural states (a state consist in
-\emph{behavioural elements} (joint states of the two interacting agents,
-associated to the current speech elements, as well as the current joint
-\emph{spatial formation}), followed by the \todo[inline]{finish that}.
+%They rely on unsupervised learning, with a
+%training stage consisting in clustering behavioural states (a state consist in
+%\emph{behavioural elements} (joint states of the two interacting agents,
+%associated to the current speech elements, as well as the current joint
+%\emph{spatial formation}), followed by the
 
@@
 are extracted and reprojected onto the 3D point-clouds (Fig.~\ref{fig|rviz},
-right). Faces have been successfully detected in 57\% of the recorded frames
+right). Faces have been successfully detected in 56\% of the recorded frames
 in the dataset. Undetected faces are due to occlusions (the child put her arm in

### 2017-06-27 paper.tex
Completed intro

@@
 These examples show the burgeoning interest of our community for the automatic
-learning of social interactions, but also highlight the lack of structure of
-these research efforts, as further illustrated by the lack of large, open
-datasets of social interactions, suitable for machine-learning applications.
+learning of social interactions, but also highlight the lack of structured
+research efforts, as further illustrated by the limited availability of large
+and open datasets of social interactions, suitable for machine-learning
+applications.
 
-One interesting dataset is the \emph{Multimodal Dyadic Behavior
-Dataset} (\emph{MMDB},~\cite{rehg2013decoding}). It comprises of 160 sessions of
-3 to 5 minute child-adult interactions. During these interactions, the
-experimenter plays with toddlers (1.5 to 2.5 years old) in semi-structured ways.
-The dataset includes video streams of the faces and the room, audio, physiological data
-(electrodermal activity) as well as manual annotations of certain behaviours
-(like gaze to the examiner, laughter, pointing). The lack of calibration and camera positioning
-information in the dataset limits the possibilities of automatically extracting and
-labelling key interaction features (like facial expressions or joint gaze). Besides,
-this dataset focuses on very young children with in adult-driven interactions,
-and as such does not include any episodes of naturally-occurring social
-interactions between peers.
+One such dataset is the \emph{Multimodal Dyadic Behavior Dataset}
+(\emph{MMDB},~\cite{rehg2013decoding}). It comprises of 160 sessions of 3 to 5
+minute child-adult interactions. During these interactions, the experimenter
+plays with toddlers (1.5 to 2.5 years old) in semi-structured ways.  The dataset
+includes video streams of the faces and the room, audio, physiological data
+(electrodermal activity) as well as manual annotations of specific behaviours
+(like gaze to the examiner, laughter, pointing). While this dataset is in
+principle relevant, it focuses on very young children during short, adult-driven
+interactions. As such, it does not include episodes of naturally-occurring
+social interactions between peers, and the diversity of said interactions is
+limited. Besides, the lack of intrinsic and extrinsic camera calibration
+information in the dataset prevent the automatic extraction and labelling of key
+interaction features (like mutual gaze).
 
-\todo[inline]{look at other child-child datasets!}
-\todo[inline]{mention research on deep-learning to diagnose autism}
+Another recent dataset, the \emph{Tower Game Dataset}~\cite{salter2015tower},
+focuses specifically on rich dyadic social interactions. The dataset comprises
+of 39 adults recorded over a total of 112 annotated sessions of 3 min in
+average. The participant are instructed to build together a tower with wooden
+blocks. Interestingly, the participants are not allowed to talk in order to
+maximise the amount of non-verbal communication. The skeletons and faces of the
+participants are recorded, and the dataset is manually annotated with so-called
+\emph{Essential Social Interaction Predicates} (ESIPs): rhythmic coupling
+(entrainment or attunement), mimicry (behavioral matching), movement
+simultaneity, kinematic turn taking patterns, joint attention. The dataset does
+not appear to be publicly available on-line.
+
+The PInSoRo dataset shares the aims of the \emph{Tower Game Dataset}, with
+however a stronger focus on natural, real-world behaviours: as presented in the
+following sections, we record much longer interactions (up to 40 minutes) of
+free-play interaction, fostering a wider range of socio-cognitive behaviours.
 
@@
 
-\todo[inline]{the dataset is not specific to robotics...PinSoRo is maybe not the
-    best name?}
-
 The Plymouth Interacting Social Robots (PInSoRo) Dataset comprises of natural
 social interactions arising between children, and recorded with machine learning
-applications in mind.
+and social robotics applications in mind.
 

### 2017-06-28 paper.tex
Update on Overleaf.

@@
 this task influences in fundamental ways the kind of interactions that might be
-observed, and thus, learned. Because they usually target very specific social
+observed, and later, learned. Because they typically target specific social
 or cognitive skills in a tightly controlled manner, traditional socio-cognitive tasks are
 usually simple, constrained tasks that do not reflect the complexity and dynamics of
-real-world interactions. On the contrary, this dataset aims at capturing a rich set
-of naturally-occurring social interactions. As such, a suitable task has to be:
+real-world interactions. Conversely, this dataset aims at capturing a rich set
+of naturally-occurring social interactions, and participants should engage in a task that:
 
 \begin{itemize}
+    \item is not directed, to maximise natural, non-contrived behaviours;
     \item elicit a large range of interaction situations;
@@
         behaviours are to be observed;
-    \item fundamentally social, \ie the task would make little sense for an
-        agent alone;
-    \item foster non-trivial social dynamics, such as implicit turn-taking.
+    \item evidence social dynamics, such as coupling, joint attention, turn-taking.
 \end{itemize}
 
-We present a new experimental task based on \emph{free play} interactions. Pairs
-of children (4-7 years old) are invited to freely draw and interact with items
-displayed on an interactive table, telling and acting stories with no explicit
-goal set in advance (Fig.~\ref{fig|freeplay}). The task is designed so that
-children can engage in open-ended and non-directive play situations, yet
-sufficiently well defined to be controlled, suitable for detailled recording,
-and easily reproducible.
+We have designed a new experimental task, called the \emph{freeplay sandbox}, and based on 
+free play interactions. Pairs of children (4-8 years old) are invited to freely draw and interact with items
+displayed on an interactive table, without any explicit goal set by the experimenter (Fig.~\ref{fig|freeplay}).
+The task is designed so that children can engage in open-ended and non-directive play, yet it is
+sufficiently defined to be suitable for recording, and reproducible in comparable conditions.
 
@@
 
-The \emph{freeplay sandbox} follows the sandtray
+The freeplay sandbox follows the sandtray
 paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
-with multitouch support) is used as an interactive surface (\emph{sandtray}). The two children play together
+with multitouch support) is used as an interactive surface (\emph{sandtray}). Two children play together
 by freely moving interactive items (Fig.~\ref{fig|sandbox}) on the surface. A background image
@@
 grass, beach, bushes...). By drawing on top of the background picture, the
-children can also alter this environment to their liking. The players do not have any particular task to
-complete, they are simply invited to freely play for as long as they wish.
+children can change the environment to their liking. The players do not have any particular task to
+complete, they are simply invited to freely play for as long as they want.
 
 In order to constraint the dataset to a tractable domain, the task is limited to
-a dyadic \emph{face-to-face} interaction.  The resulting interaction space is
+a dyadic face-to-face interaction.  The resulting interaction space is
 nevertheless high-dimensional, with two main axis: the game-related actions
@@
 (monitoring the other's actions, discussing and agreeing on joint actions,
-helping behaviours, etc.).
+helping/hostile behaviours, etc.).
 
@@
 The dataset consists in a collection of records. Each record correspond to one
-play interaction between two children. To date (June 2017) 25 records have been acquired (\ie 50 children). At the end of the acquisition campaign
+play interaction between two children. To date (June 2017) 25 records have been
+acquired (\ie 50 children), totalling 08h41m of recording. At the end of the acquisition campaign
 (July 2017), the dataset is planned to include 50 records. As
@@
             & face (depth)                          & VGA (640$\times$480), 30Hz           \\
-            & facial features                       & 68 3D points (point cloud)           \\ \midrule
+            & facial features                       & 68 3D points, 30Hz                   \\ \midrule
 child 2     & audio                                 & 16kHz, mono, semi-directional        \\
@@
             & face (depth)                          & VGA (640$\times$480), 30Hz           \\
-            & facial features                       & 68 3D points (point cloud)           \\ \midrule
+            & facial features                       & 68 3D points, 30Hz                   \\ \midrule
 environment & RGB                                   & qHD (960$\times$540), 29.7Hz         \\ \midrule
 touchscreen & background drawing (RGB)              & 4Hz                                  \\
-            & touches                               & 6 points multi-touch                 \\
-            & items position and orientation        & ($x,y,\theta$)                       \\ \midrule
-others      & \multicolumn{2}{l}{annotations of social behaviours and remarkable events}   \\
-            & \multicolumn{2}{l}{static transforms between touchscreen and facial cameras} \\
+            & touches                               & 6 points multi-touch, 10Hz           \\
+            & items position and orientation        & ($x,y,\theta$), 10Hz                 \\ \midrule
+annotations & \multicolumn{2}{l}{timestamped annotations of social behaviours and remarkable events}   \\ \midrule
+other       & \multicolumn{2}{l}{static transforms between touchscreen and facial cameras} \\
             & \multicolumn{2}{l}{cameras calibration informations}                         \\ \bottomrule
@@
 
-\subsection{Hardware and Software}
+\subsection{Apparatus}
 
@@
 
-While
-the raw RGB and depth streams are neither rectified nor registered together (due
+The dataset contains the raw datastreams, as recorded during the study. No post-processing has been applied.
+% in the robot condition, the interaction bag is trimmed to start once the robot is localised
+While the raw RGB and depth video streams are neither rectified nor registered together (due
 to performance considerations during the recording), it can be easily achieved

### 2017-06-28 paper.tex
Update on Overleaf.

@@
 Child-robot interactions are increasingly being explored in domains which
-require longer-term application, such as healthcare and education. In order
-for a robot to behave in an appropriate manner over longer timescales, its
+require longer-term applications, such as healthcare and education. In order
+for a robot to behave in an appropriate manner over longer time scales, its
 behaviours should be aligned and meaningful to the interacting children.
@@
 so, we capture a rich set of behavioural patterns occurring in natural
-social interactions. Learning to recognise, and possibly generate, such
-behaviours is one of the future research direction that we foresee this dataset to effectively support.
+social interactions between children. Learning to recognise, and possibly generate, such
+behaviours is one of the future research directions that we foresee this dataset to effectively support.
 
@@
 While the family of recurrent neural networks (RNN) have repeatedly made the
-headlines over the last few years with impressive results, notably in image
-classification, image labelling and automatic translation, they have been so far 
+headlines over the last few years with impressive results---notably in image
+classification, image labelling and automatic translation---they have been so far 
 largely ignored by  many other fields as they are perceived to require very
 large datasets (hundreds of thousands to millions of observations) to actually
-build up useful capabilities. Even though neural networks have demonstrated
+build up useful capabilities. 
+%CERE: perhaps this sentence could be shorter?
+Even though neural networks have demonstrated
 compelling results in open-ended, under-specified tasks like image labelling, they
@@
 interactions.
-
 Besides, if one considers ``social interactions'' to also entail joint
@@
 
-In 2015, the Google DeepMind team demonstrated how a convolutional
+That being said, in 2015, the Google DeepMind team demonstrated how a convolutional
 recurrent neural network could learn to play the game Break-Out (amongst
 48 other Atari games) by only looking at the gaming console
-screen~\cite{mnih2015human}. This result represents a major milestone: they show
+screen~\cite{mnih2015human}. This result represents a major milestone: they showede
 that a relatively small sample size (about 500 games) is sufficient for an 

### 2017-06-28 paper.tex
Update on Overleaf.

@@
 48 other Atari games) by only looking at the gaming console
-screen~\cite{mnih2015human}. This result represents a major milestone: they showede
+screen~\cite{mnih2015human}. This result represents a major milestone: they showed
 that a relatively small sample size (about 500 games) is sufficient for an 

### 2017-06-28 paper.tex
Polishing

@@
 \usepackage{booktabs}
+\usepackage{tablefootnote}
 
@@
 
-\subsection{The Freeplay Sandbox}
+\subsection{The Freeplay Sandbox task}
 
@@
 (July 2017), the dataset is planned to include 50 records. As
-mentioned above, the duration of each play episode varies ($M=20m51s,
-SD=10m40s$), but it is capped to a maximum of 40 minutes.
+the children decide themselves when to stop, the duration of each play episode varies ($M=20m51s,
+SD=10m40s$). It is however capped to a maximum of 40 minutes.
 
 Data is collected using the ROS middleware\footnote{\url{http://ros.org}} and
-recorded as \emph{bag} files. Every dataframe is timestamped.
-Table~\ref{table|datastreams} lists all the recorded datastreams.
+recorded as \emph{bag} files. Table~\ref{table|datastreams} lists all the
+recorded datastreams.  Every dataframe is timestamped; as the data is recorded
+using ROS's bag files, it can be replayed in the exact same conditions as it
+has been recorded.  All the video streams use calibrated cameras; only the raw
+RGB and depth video streams are stored in the dataset, section~\ref{postprocessing} discuss
+how to reconstruct from them calibrated streams and 3D point clouds.
 
@@
             & face (depth)                          & VGA (640$\times$480), 30Hz           \\
-            & facial features                       & 68 3D points, 30Hz                   \\ \midrule
+            & facial features\tablefootnote{\label{postprocesspass}Added to the dataset during a post-processing pass.} & 68 3D points, 30Hz                   \\ \midrule
 child 2     & audio                                 & 16kHz, mono, semi-directional        \\
@@
             & face (depth)                          & VGA (640$\times$480), 30Hz           \\
-            & facial features                       & 68 3D points, 30Hz                   \\ \midrule
+            & facial features\textsuperscript{\ref{postprocesspass}} & 68 3D points, 30Hz                   \\ \midrule
 environment & RGB                                   & qHD (960$\times$540), 29.7Hz         \\ \midrule
@@
 
-As the data is recorded using ROS's bag files, it can be replayed in the exact
-same conditions as it was during the recording. All the video streams use
-calibrated cameras. Only the raw RGB and depth video streams are recorded.
-
 \subsection{Apparatus}
@@
 The sandtray is made of a 27" Samsung All-In-One computer running Ubuntu Linux
-and equipped with a fast 1TB SSD hard-drive. The computer is held horizontal in
+and equipped with a fast 1TB SSD hard-drive. The computer is held horizontally in
 a custom aluminium frame standing 26cm above the ground
-(Fig.~\ref{fig|freeplay}, right). All the cameras are plugged directly over
+(visible in Fig.~\ref{fig|freeplay}). All the cameras are plugged directly over
 USB-3 to the computer. The computer performs all the data acquisition using ROS
@@
 
-The children faces are recorded from two Intel RealSense SR300 RGB-D cameras
-(short range, 0.2m to 1.2m, indoor RGB-D cameras), placed at the corners of the
-the sandtray (Figs.~\ref{fig|rviz} and~\ref{fig|setup}), and tilted to face the
-children. The cameras are rigidly mounted on custom 3D-printed brackets, and
-their 6D pose with respect to the sandtray screen is precisely known and
-recorded (this enables in particular accurate estimation of the children's head
-pose).
+The children faces are recorded from two short range Intel RealSense SR300
+RGB-D cameras (0.2m to 1.2m) placed at the corners of the the sandtray
+(Figs.~\ref{fig|rviz} and~\ref{fig|setup}) and tilted to face the children. The
+cameras are rigidly mounted on custom 3D-printed brackets. This enables a
+precise measurement of their 6D pose relative to the touchscreen (extrinsic
+calibration).
 
 Audio is recorded from the same SR300 cameras (one audio stream is recorded for
-each child, from the camera facing her/him).
+each child, from the camera facing her).
 
@@
 the whole interaction setting. These video stream is intended to support human
-coders while annontating the interaction, and is not precisely calibrated.
+coders while annotating the interaction, and is not precisely calibrated.
 
@@
     \caption{The freeplay sandbox, visualised at runtime with ROS RViz. The
-    current background drawing is published on a regular image
-    topic and computer vision is used to segment it into zones (visible in
-    the central panel, right picture). The poses of the interactive items
-    are published, as well as an estimate of the children's head pose.
-    The left picture shows a reconstructed 3D point cloud of one child face,
-    along with the detected 68 facial features.}
+    poses of the interactive items are published (visible in the central panel,
+    right picture), as well as an estimate of the children's head pose.  The
+    left picture shows a reconstructed 3D point cloud of one child face, along
+    with the detected 68 facial features.}
+
     \label{fig|rviz}
@@
 \paragraph{Software}
-The sandbox is implemented with two frameworks: the Qt's \emph{QtQuick} framework
+The sandbox is implemented with two frameworks: the Qt's QtQuick framework
 for the graphical interface of the game, and the \emph{Robot Operating System}
@@
 \subsection{Data post-processing}
+\label{postprocessing}
 

### 2017-06-28 paper.tex
Demographics

@@
 pipelines. A dedicated bridge between QtQuick and ROS has been specifically
-developped to enable the game interface to export the positions of every
-iteractive items as they move, the background image whenever it is painted over,
+developed to enable the game interface to export the positions of every
+interactive items as they move, the background image whenever it is painted over,
 and the children's touches. The game interface is open-source and available online:
@@
 
-Finally, we have developped a web-based \emph{supervisor} that permits the
+Finally, we have developed a web-based \emph{supervisor} that enables the
 experimenter to remotely start/stop the ROS nodes and the game GUI, as well as to
 record annotations during the experiment. The supervisor ensures that the exact
-same recording procedure (detailled in the next section) is followed for every
+same recording procedure (detailed in the next section) is followed for every
 participants. The supervisor is available online as well:
@@
 \subsection{Dataset Acquisition}
+
+The experimental setup is presented in Fig.~\ref{fig|setup}.  The experimenter
+stays in the room, visible to the child, during the whole experiment. When
+requested by the nursery or school, another childminder (nursery/school staff)
+can be present in the room. She/he is however asked not to step in during the
+experiment.
 
@@
 
-The experimenter stays in the room, visible to the child. When
-requested by the nursery or school, another childminder (nursery/school
-staff) can be present in the room. She/he is however asked not to step
-in during the experiment.
+\paragraph{Demographics}
 
-2 age groups, balanced across conditions: 4 y.o. (nursery) and 6 to 7
-y.o. (Y1/Y2)
-
-Normally developing children, from local primary schools and nurseries.
+We recruited normally developing children, from local primary schools and
+nurseries. The children are aged 4 to 8 ($M=5.8$). So far, 50 children (23
+female, 27 males) were recorded in pairs. Pairs were formed by the school or
+nursery staff, and all the pairs knew each other before hand.
 
@@
 \toprule
-\bf Welcome                       & \tabitem explain the purpose of the study: showing robots how children play  \\
+\bf Greetings                     & \tabitem explain the purpose of the study: showing robots how children play  \\
 \emph{(about 5 min)}              & \tabitem briefly present a Nao robot -- the robot stands up, gives a short

### 2017-06-28 paper.tex
Update on Overleaf.

@@
 
-While the family of recurrent neural networks (RNN) have repeatedly made the
+While the family of recurrent neural networks (RNNs) have repeatedly made the
 headlines over the last few years with impressive results---notably in image
 classification, image labelling and automatic translation---they have been so far 
+%CERE: do these need references?
 largely ignored by  many other fields as they are perceived to require very
@@
 12 DoF of the two arms. Successfully folding towels entails an explicit sequencing of
-actions (therefore implicit temporal modeling): the fact that such a complex
+%CERE: do you need to define DoF or is it obvious?
+actions (therefore implicit temporal modeling). The fact that such a complex
 process can be successfully learned from a small training dataset should lead
@@
 that the complexity of mechanisms that such neural networks have been able to
-uncover and model should invite our community to question its applicability to
-human-robot interactions in general, and sustained, natural child-robot
-interactions in particular. The current lack of appropriate HRI
-datasets suitable for neural networks training is however an obstacle: we
-introduce in this article the PInSoRo dataset, a large dataset of child-child
+uncover and model should invite our community to explore its applicability to
+%CERE: ``question'' to me seemed negative
+human-robot interactions (HRI) in general, and sustained, natural child-robot
+interactions in particular. However, the current lack of appropriate HRI
+datasets suitable for neural networks training is an obstacle. In this article we introduce the PInSoRo dataset, a large dataset of child-child
 natural interactions suitable for machine learning.
@@
 previously explored, and can be considered as the extension of the traditional
-learning from demonstration (LfD) paradigms to social interactions (for
-instance~\cite{nehaniv2007imitation,mohammad2015interaction}). Existing
-research focuses however on low-level recognition or generation of short,
+learning from demonstration (LfD) paradigms to social interactions (i.e.,~\cite{nehaniv2007imitation,mohammad2015interaction}). However, existing
+research focuses on low-level identification or generation of short,
+%CERE: ``recognition'' confused me as it's a type of memory paradigm -- possibly that's irrelevant to this audience though
 self-standing behaviours, including social gestures~\cite{nagai2005learning}
@@
 applications.
+%CERE: you've already said most of this, perhaps it could be cut?
 
@@
 of 39 adults recorded over a total of 112 annotated sessions of 3 min in
-average. The participant are instructed to build together a tower with wooden
-blocks. Interestingly, the participants are not allowed to talk in order to
+average. The participants are instructed to build a tower together with wooden
+blocks. Interestingly, the participants are not allowed to talk to
 maximise the amount of non-verbal communication. The skeletons and faces of the
@@
 
-The Plymouth Interacting Social Robots (PInSoRo) Dataset comprises of natural
-social interactions arising between children, and recorded with machine learning
+The PInSoRo Dataset comprises of natural
+social interactions arising between children, recorded with machine learning
 and social robotics applications in mind. 
-
 As detailed in section~\ref{availability}, this dataset is open and made freely
@@
 of naturally-occurring social interactions, and participants should engage in a task that:
+%CERE: this last sentence read a little oddly to me... not quite sure why to fix it though
 
@@
 
-We have designed a new experimental task, called the \emph{freeplay sandbox}, and based on 
+We have designed a new experimental task, called the \emph{freeplay sandbox}, that is based on 
 free play interactions. Pairs of children (4-8 years old) are invited to freely draw and interact with items
@@
 a dyadic face-to-face interaction.  The resulting interaction space is
-nevertheless high-dimensional, with two main axis: the game-related actions
+nevertheless high-dimensional, with two main axes: the game-related actions
 (drawing, manipulating items, telling stories) and the social interactions
@@
 
-The dataset consists in a collection of records. Each record correspond to one
+The dataset consists of a collection of records. Each record correspond to one
 play interaction between two children. To date (June 2017) 25 records have been
@@
 has been recorded.  All the video streams use calibrated cameras; only the raw
-RGB and depth video streams are stored in the dataset, section~\ref{postprocessing} discuss
+RGB and depth video streams are stored in the dataset, Section~\ref{postprocessing} discusses
 how to reconstruct from them calibrated streams and 3D point clouds.
@@
     left picture shows a reconstructed 3D point cloud of one child face, along
+    %CERE: left? central? this caption confused me a bit... 
     with the detected 68 facial features.}
@@
                                     play and soon I'll be able to play with you. Is that alright?"}), and
-                                    sit down. The robot is then put aside and does not intervene anymore
+                                    sits down. The robot is then put aside and does not intervene
                                     for the remaining of the experiment; \\
@@
 right). Faces have been successfully detected in 56\% of the recorded frames
-in the dataset. Undetected faces are due to occlusions (the child put her arm in
+in the dataset. Undetected faces are due to occlusions (a child put her arm in
 front of the camera), large head rotations or quick child motion.
@@
 \bf Construct          & \bf Description                                                                                                                                                                                                                                                                                                                                           \\ \midrule
-Hostile/coercive   & anti-social behavior ; examples include grabbing, hitting, taunting, provoking, yelling.                                                       \\
-Prosocial/engaged  & engaged in a positive social interaction ; examples include on-task work or conversations, helping, sharing, and compromising.                 \\
-Assertive          & child stating a position or opinion ; examples include asserting beliefs (verbally or nonverbally), directing the partner.                     \\
-Passive/disengaged & lack of involvement in the, activity, physical disengagement ; examples include staring into space, putting head down, speaking quite softly, and walking away from an activity.\\
-Adult seeking      & effort to seek assistance or attention from an adult ; examples include requesting help.                                                       \\
-Irrelevant         & child being actively engaged in an off-task behavior ; examples include starting a different activity, fiddling with objects as a main focus of attention, goofy or silly behaviors.
+Hostile/coercive   & anti-social behavior; examples include grabbing, hitting, taunting, provoking, yelling.                                                       \\
+Prosocial/engaged  & engaged in a positive social interaction; examples include on-task work or conversations, helping, sharing, and compromising.                 \\
+Assertive          & child stating a position or opinion; examples include asserting beliefs (verbally or nonverbally), directing the partner.                     \\
+Passive/disengaged & lack of involvement in the activity, physical disengagement; examples include staring into space, putting head down, speaking quite softly, and walking away from an activity.\\
+Adult seeking      & effort to seek assistance or attention from an adult; examples include requesting help.                                                       \\
+Irrelevant         & child being actively engaged in an off-task behavior; examples include starting a different activity, fiddling with objects as a main focus of attention, goofy or silly behaviors.
 \\ \bottomrule

### 2017-06-28 paper.tex
Moving around figures

@@
     \centering
-    \includegraphics[width=0.55\linewidth]{setup-child-child.png}
+    \includegraphics[width=0.55\linewidth]{setup_child_child_top}
     \hspace{1em}
     \includegraphics[width=0.4\linewidth]{child-child-env}
-    \caption{The free play social interactions sandbox: two children interact in
-    a free-play situation, by drawing and manipulating items on a touchscreen.}
+    \caption{The free-play social interactions sandbox: two children interact in
+    a free-play situation, by drawing and manipulating items on a touchscreen.
+    Children are facing each other and sit on cushions. Each child wears a
+    bright sport bib, either purple or yellow, to facilitate later
+    identification.}
+
     \label{fig|freeplay}
@@
 
-The experimental setup is presented in Fig.~\ref{fig|setup}.  The experimenter
+The experimental setup is visible in Fig.~\ref{fig|freeplay}, left.  The experimenter
 stays in the room, visible to the child, during the whole experiment. When
@@
 experiment.
-
-\begin{figure}[htbp]
-    \centering
-    \includegraphics[width=0.8\linewidth]{setup_child_child_top}
-    %\hspace{1.5cm}
-    %\includegraphics[width=0.4\linewidth]{setup_child_robot_top}
-    %\caption{Experimental setup, child-child condition (left) and child-robot condition (right).}
-    \caption{Experimental setup. Children are facing each other and sit on
-    cushions. Each child wears a bright sport bib, either purple or yellow, to
-    facilitate later identification.}
-    \label{fig|setup}
-\end{figure}
 
@@
 
-The following acquisition procedure has been followed with all participants.
+The following acquisition procedure has been followed with all participants. To
+ease later identification, each child was given a colourful (purple or yellow)
+sport bib to wear.
 
@@
 \bf Greetings                     & \tabitem explain the purpose of the study: showing robots how children play  \\
-\emph{(about 5 min)}              & \tabitem briefly present a Nao robot -- the robot stands up, gives a short
+\emph{(about 5 min)}              & \tabitem briefly present a Nao robot: the robot stands up, gives a short
                                     message (\emph{"Hello, I'm Nao. Today, I'll be watching you while you
@@
                                     sits down. The robot is then put aside and does not intervene
-                                    for the remaining of the experiment; \\
+                                    for the remaining of the experiment \\
                                     %Condition B: "Hello, I'm Nao. Today I'll be playing with you. Exciting!")\\ 
@@
                                   & \tabitem remind the children that they can withdraw at anytime \\ \midrule
-\bf Visual focus task             &  \\ 
+        \bf Visual focus task             &  \emph{}\\ 
 \emph{(30 sec)}                   &  \\ \midrule
@@
 computed and coded (mutual gaze, for instance), some are human-coded.
-
 The annotations are integrated in the dataset as a timestamped datastream.
@@
 Olswang~\etal\cite{olswang2006reliability}. It consists in 6 constructs
-characterising social communication (table~\ref{table|sccs}).
+characterising social communication (Table~\ref{table|sccs}).
 
@@
 \end{table}
+
+%CERE: this didn't table for me?
 
@@
 
-The dataset is build with machine learning applications in mind, and
+The dataset is built with machine learning applications in mind, and
 particularly, applications to socio-cognitive robotics.
 
-Finally, the social setting (two players facing each other) implicitely invites
+Finally, the social setting (two players facing each other) implicitly invites
 social interaction: onlooker play, for instance, is unlikely to be observed as

### 2017-06-28 paper.tex
Moving around pictures

--- 2017-06-28 paper.tex
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{sandbox}
+    \includegraphics[width=0.52\linewidth]{sandbox}
+    \hspace{0.2em}
+    \includegraphics[width=0.46\linewidth]{rviz-sandtray}
     \caption{Example of a possible game situation. Items (animals,
     characters...) can be dragged over the whole play area, while the background
-    picture can be painted over by picking a colour.}
+    picture can be painted over by picking a colour. The right picture shows a
+    real-time visualisation of the experiment using ROS RViz.}
 
@@
     \centering
-    \includegraphics[width=0.5\linewidth]{rviz-sandtray}
-    \hspace{0.2em}
-    \includegraphics[width=0.48\linewidth]{3d-point-cloud-facial-features}
-    \caption{The freeplay sandbox, visualised at runtime with ROS RViz. The
-    poses of the interactive items are published (visible in the central panel,
-    right picture), as well as an estimate of the children's head pose.  The
-    left picture shows a reconstructed 3D point cloud of one child face, along
-    %CERE: left? central? this caption confused me a bit... 
-    with the detected 68 facial features.}
+    \includegraphics[width=0.9\linewidth]{3d-point-cloud-facial-features}
+    \caption{The reconstructed 3D point cloud of one child face with the
+    detected 68 facial features, visualised in RViz.}
 
@@
 By relying on ROS for the data acquisition, real-time monitoring of the
-interaction is also possible (Fig.~\ref{fig|rviz}).
+interaction is also possible (Fig.~\ref{fig|sandbox}, right).
 The ROS data acquisition pipeline is open-source as well, and available from
@@
 republish registered streams and the corresponding 3D point-clouds (as seen in
-Fig.~\ref{fig|rviz}, left).
+Fig.~\ref{fig|rviz}).
 
@@
 extraction.}~\cite{lemaignan2016realtime}, children faces and facial features can
-are extracted and reprojected onto the 3D point-clouds (Fig.~\ref{fig|rviz},
-right). Faces have been successfully detected in 56\% of the recorded frames
+are extracted and reprojected onto the 3D point-clouds (Fig.~\ref{fig|rviz}). Faces have been successfully detected in 56\% of the recorded frames
 in the dataset. Undetected faces are due to occlusions (a child put her arm in
@@
 
-% Please add the following required packages to your document preamble:
-% \usepackage{booktabs}
 \begin{table}[]

### 2017-06-28 paper.tex
Addressed Charlotte's comments

@@
 
-While the family of recurrent neural networks (RNNs) have repeatedly made the
+The family of recurrent neural networks (RNNs) have repeatedly made the
 headlines over the last few years with impressive results---notably in image
-classification, image labelling and automatic translation---they have been so far 
-%CERE: do these need references?
-largely ignored by  many other fields as they are perceived to require very
-large datasets (hundreds of thousands to millions of observations) to actually
-build up useful capabilities. 
-%CERE: perhaps this sentence could be shorter?
-Even though neural networks have demonstrated
-compelling results in open-ended, under-specified tasks like image labelling, they
-did not stand out as attractive approaches to problems involving high dimensions
-with relatively small datasets available -- like human-robot social
-interactions.
-Besides, if one considers ``social interactions'' to also entail joint
-behavioural dynamics, and therefore, some sort of temporal modeling, neural
-networks look even less enticing as time is notably absent from most of the
-tasks which neural networks have been successful at.
+classification, image labelling and automatic translation. They have been
+however largely ignored by other fields as they are perceived to require
+impossibly large datasets (hundreds of thousands to millions of observations) to
+actually build up useful capabilities.  Even though neural networks have
+demonstrated compelling results in open-ended, under-specified tasks like image
+labelling, they did not stand out as attractive approaches to problems involving
+high dimensions with relatively small datasets available -- like human-robot
+social interactions.  Besides, if one considers ``social interactions'' to also
+entail joint behavioural dynamics, and therefore, some sort of temporal
+modeling, neural networks look even less enticing as time is notably absent from
+most of the tasks which neural networks have been successful at.
 
@@
 sequences. The network inputs are the raw video stream from the head camera and the
-12 DoF of the two arms. Successfully folding towels entails an explicit sequencing of
-%CERE: do you need to define DoF or is it obvious?
+12 degrees of freedom of the two arms. Successfully folding towels entails an explicit sequencing of
 actions (therefore implicit temporal modeling). The fact that such a complex
@@
 uncover and model should invite our community to explore its applicability to
-%CERE: ``question'' to me seemed negative
 human-robot interactions (HRI) in general, and sustained, natural child-robot
-interactions in particular. However, the current lack of appropriate HRI
-datasets suitable for neural networks training is an obstacle. In this article we introduce the PInSoRo dataset, a large dataset of child-child
-natural interactions suitable for machine learning.
+interactions in particular.
+%However, the current lack of appropriate HRI
+%datasets suitable for neural networks training is an obstacle. In this article
+%we introduce the PInSoRo dataset, a large dataset of child-child natural
+%interactions suitable for machine learning.
 
@@
 previously explored, and can be considered as the extension of the traditional
-learning from demonstration (LfD) paradigms to social interactions (i.e.,~\cite{nehaniv2007imitation,mohammad2015interaction}). However, existing
+learning from demonstration (LfD) paradigms to social interactions
+(i.e.,~\cite{nehaniv2007imitation,mohammad2015interaction}). However, existing
 research focuses on low-level identification or generation of short,
-%CERE: ``recognition'' confused me as it's a type of memory paradigm -- possibly that's irrelevant to this audience though
-self-standing behaviours, including social gestures~\cite{nagai2005learning}
-and gazing behaviours~\cite{calinon2006teaching}.
+self-standing behaviours, including social gestures~\cite{nagai2005learning} and
+gazing behaviours~\cite{calinon2006teaching}.
 
@@
 applications.
-%CERE: you've already said most of this, perhaps it could be cut?
 
@@
 
-%CERE: this didn't table for me?
-
 \begin{figure}

### 2017-06-28 paper.tex
WIP

@@
 
-\title{PInSoRo: A Dataset for Deep-learning of Human-Robot Social Interactions}
+\title{PInSoRo: A Dataset for Deep-Learning of Human-Robot Social Interactions}
 
@@
     \caption{The reconstructed 3D point cloud of one child face with the
-    detected 68 facial features, visualised in RViz.}
+    68 detected facial features, visualised in RViz.}
 
@@
 
-The experimental setup is visible in Fig.~\ref{fig|freeplay}, left.  The experimenter
-stays in the room, visible to the child, during the whole experiment. When
-requested by the nursery or school, another childminder (nursery/school staff)
-can be present in the room. She/he is however asked not to step in during the
-experiment.
+The experimental setup is visible in Fig.~\ref{fig|freeplay}, left.  The
+experimenter stays in the room, visible to the child, during the whole
+experiment. When requested by the nursery or school, another childminder
+(nursery/school staff) can be present in the room. She/he is however asked not
+to step in during the experiment. Not pictured on Fig.~\ref{fig|freeplay} is an
+'ideas box', containing 5 cards depicting simple environments (zoo, forest,
+city, space, pirates).  Children are invited by the experimenter to pick up a
+card whenever they run out of activity ideas.
 
@@
 female, 27 males) were recorded in pairs. Pairs were formed by the school or
-nursery staff, and all the pairs knew each other before hand.
+nursery staff, and all the pairs knew each other beforehand.
 
@@
                                   & \tabitem remind the children that they can withdraw at anytime \\ \midrule
-        \bf Visual focus task             &  \emph{}\\ 
+\bf Visual focus task             & \emph{(unrelated visual task where children gaze at a moving target)}\\ 
 \emph{(30 sec)}                   & \\ \midrule
@@
 \emph{(1-2 min)}                  & \\ \midrule
-\bf Item placement game           &  \\
+\bf Items placement game          & \emph{(unrelated task on spatial reasoning -- as this task uses the same graphical environment as the main free-play task, it helps the children to familiarise themselves with the game environment)} \\
 \emph{(5 to 10 min)}              &  \\ \midrule
-\bf Freeplay task                 & \tabitem start the recording \\
-\emph{(up to 40 min)}             & \tabitem initial prompt: \emph{"Just to remind you, you can use the animals or draw. Whatever you
+\bf Freeplay task                 & \tabitem initial prompt: \emph{"Just to remind you, you can use the animals or draw. Whatever you
                                   like. If you run out of ideas, there's also an ideas box. For example, the first one is a
@@
 
+No rewards were given to the children beyond a sticker and, at the end of the
+acquisition campaign, a demonstration of robots with the whole class.
+
 \subsection{Data post-processing}
@@
 
-The dataset contains the raw datastreams, as recorded during the study. No post-processing has been applied.
+The dataset contains the raw datastreams, as recorded during the study.
 % in the robot condition, the interaction bag is trimmed to start once the robot is localised
-While the raw RGB and depth video streams are neither rectified nor registered together (due
-to performance considerations during the recording), it can be easily achieved
-as a post-process step, using standard ROS tools (the ROS {\tt rgbd\_pipeline}).
-Pre-configured scripts are available alongside the dataset\footnote{Available
-online from
+While the raw RGB and depth video streams are neither rectified nor registered
+together (due to performance considerations during the recording), it can be
+easily achieved using standard ROS tools (the ROS {\tt
+rgbd\_pipeline}). Pre-configured scripts are available alongside the
+dataset\footnote{Available online from
 \url{https://github.com/severin-lemaignan/freeplay-sandbox-analysis}.} to
@@
 dlib} face tracker for facial features
-extraction.}~\cite{lemaignan2016realtime}, children faces and facial features can
-are extracted and reprojected onto the 3D point-clouds (Fig.~\ref{fig|rviz}). Faces have been successfully detected in 56\% of the recorded frames
-in the dataset. Undetected faces are due to occlusions (a child put her arm in
-front of the camera), large head rotations or quick child motion.
+extraction.}~\cite{lemaignan2016realtime}, children faces and facial features
+can be extracted and reprojected onto the 3D point-clouds
+(Fig.~\ref{fig|rviz}). Faces have been successfully detected in 56\% of the
+recorded frames in the dataset -- undetected faces are due to occlusions (a child
+put her arm in front of the camera), large head rotations or quick child
+motion -- and whenever a face is present, the corresponding 3D facial features
+have been added to the dataset.
 
@@
 
-The last part of the dataset consist in annotations of the
-children's social behaviours. Some of these behaviours are automatically
-computed and coded (mutual gaze, for instance), some are human-coded.
-The annotations are integrated in the dataset as a timestamped datastream.
+The last part of the dataset consist in annotations of the children's social
+behaviours. These behaviours are either automatically computed and coded
+(mutual gaze, for instance), or are human-coded. The annotations are all
+timestamped and integrated as an independent datastream in the dataset.
 
@@
 
-We use the \emph{Social Communication Coding System} proposed by
-Olswang~\etal\cite{olswang2006reliability}. It consists in 6 constructs
-characterising social communication (Table~\ref{table|sccs}).
+The dataset includes two types of manual annotations: \emph{remarkable events},
+annotated by the experimenter during the experiment as free text annotation, and
+\emph{social behaviours} coded with the \emph{Social Communication Coding
+System} (SCCS) proposed by Olswang~\etal\cite{olswang2006reliability}.
+These annotations are performed post-hoc, with a dedicated tool, able to replay synchronously
+the main datastreams (Fig.~\ref{fig|analysis}).
+
+The SCCS consists in 6 mutually exclusive constructs characterising social
+communication (Table~\ref{table|sccs}). These behaviours represent a typical
+classification output that a network could be trained to recognise.
 
@@
 Prosocial/engaged  & engaged in a positive social interaction; examples include on-task work or conversations, helping, sharing, and compromising.                 \\
-Assertive          & child stating a position or opinion; examples include asserting beliefs (verbally or nonverbally), directing the partner.                     \\
+Assertive          & child stating a position or opinion; examples include asserting beliefs (verbally or non-verbally), directing the partner.                     \\
 Passive/disengaged & lack of involvement in the activity, physical disengagement; examples include staring into space, putting head down, speaking quite softly, and walking away from an activity.\\
@@
 
-\begin{figure}
+\begin{figure}[b!]
     \centering
     \includegraphics[width=0.8\linewidth]{analysis}
-    \caption{Screenshot of the interaction analysis tool: key record information
-    are played back in a synchronised (faces and audio of the two children,
-    interaction occuring on the interactive table, etc.) while the human coder
+    \caption{Screenshot of our interaction annotation tool: the recorded interaction data
+    is played back synchronously (faces and audio of the two children,
+    interaction occurring on the interactive table, etc.) while the human coder
     annotate the record timeline with social situations of interest.}
@@
 
-Several metrics and coding scheme
-
-analysis of behavioural alignment between partners (via
-metrics like the recently proposed \emph{Individual Motor
-Signature}~\cite{slowinski2016dynamic})
-
-Słowinski’s Individual Motor Signature
-(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
-interactions,
+%Several metrics and coding scheme
+%
+%analysis of behavioural alignment between partners (via
+%metrics like the recently proposed \emph{Individual Motor
+%Signature}~\cite{slowinski2016dynamic})
+%
+%Słowinski’s Individual Motor Signature
+%(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
+%interactions,
 

### 2017-06-28 paper.tex
Added a link to the temporary website

@@
 The dataset is freely available to any interested researcher\footnote{Note to
-the reviewers: the link to the dataset's dedicated website will be added to the
-final version of the article.}. Due to ethical and data protection regulations,
-the dataset is however made available in two forms: a public, Creative-Commons
-licensed, version that do not include any video material of the children (no
-video streams, audio included); and a complete version that includes all video
-streams. This second version is freely available as well, but interested
-researchers must first fill a data protection form.
+the reviewers: for the purpose of the review, an excerpt of the dataset is
+available online: \url{https://freeplay-sandbox.github.io/}. A link to the
+complete dataset will be made available in the final version of the article.}.
+Due to ethical and data protection regulations, the dataset is however made
+available in two forms: a public, Creative-Commons licensed, version that do
+not include any video material of the children (no video streams, audio
+included); and a complete version that includes all video streams. This second
+version is freely available as well, but interested researchers must first fill
+a data protection form.
 

### 2017-06-28 paper.tex
Update on Overleaf.

@@
 Child-robot interactions are increasingly being explored in domains which
-require longer-term applications, such as healthcare and education. In order
+require longer-term interactions, such as healthcare and education. In order
 for a robot to behave in an appropriate manner over longer time scales, its
-behaviours should be aligned and meaningful to the interacting children.
-Generating such sustained and engaging social behaviours is an on-going research
+behaviours should be contingent and meaningful to the unfolding interaction.
+Generating sustained and engaging social behaviours is an on-going research
 challenge, and we argue here that the recent progress of deep machine learning
@@
 As an initial step in that direction, we introduce a novel,
-open dataset of children social interactions, designed with
+open dataset of child social interactions, designed with
 machine learning applications in mind. Our data acquisition methodology relies on
-a engaging and purposefully underspecified \emph{free play} interaction. By doing
+an engaging and purposefully underspecified \emph{free play} interaction. By doing
 so, we capture a rich set of behavioural patterns occurring in natural
@@
 
-The family of recurrent neural networks (RNNs) have repeatedly made the
-headlines over the last few years with impressive results---notably in image
+The broad family of deep neural networks have repeatedly made the
+headlines in the last few years with reports of impressive results---notably in image
 classification, image labelling and automatic translation. They have been
@@
 actually build up useful capabilities.  Even though neural networks have
-demonstrated compelling results in open-ended, under-specified tasks like image
+demonstrated compelling results in non-trivial tasks like image
 labelling, they did not stand out as attractive approaches to problems involving
-high dimensions with relatively small datasets available -- like human-robot
-social interactions.  Besides, if one considers ``social interactions'' to also
+high input and output dimensions where only relatively small datasets are available -- like human-robot
+interactions.  Furthermore, if one considers ``social interactions'' to also
 entail joint behavioural dynamics, and therefore, some sort of temporal
@@
 strategies that look like they would necessitate planning (the system first
-breaks bricks on one side to eventually get the ball to break-out and reach out the area
+breaks bricks on one side to eventually get the ball to break through and reach the area
 above the remaining bricks, therefore ensuring rapid progress in the
@@
 objects like towels using a dual-arm mobile manipulator) from only \emph{35}
-demonstrations of $\approx$ 70 seconds-long teleoperated
+demonstrations of $\approx$ 70 second-long teleoperated
 sequences. The network inputs are the raw video stream from the head camera and the
@@
 
-Indeed, as we have previously argued~\cite{lemaignan2016towards}, we believe
+%Indeed, as we have previously argued~\cite{lemaignan2016towards}, 
+We believe
 that the complexity of mechanisms that such neural networks have been able to
@@
 (i.e.,~\cite{nehaniv2007imitation,mohammad2015interaction}). However, existing
-research focuses on low-level identification or generation of short,
-self-standing behaviours, including social gestures~\cite{nagai2005learning} and
+research focuses on low-level identification or generation of brief, autonomous behaviours, including social gestures~\cite{nagai2005learning} and
 gazing behaviours~\cite{calinon2006teaching}.
@@
 These examples show the burgeoning interest of our community for the automatic
-learning of social interactions, but also highlight the lack of structured
+learning of social responses, but also highlight the lack of structured
 research efforts, as further illustrated by the limited availability of large
@@
 minute child-adult interactions. During these interactions, the experimenter
-plays with toddlers (1.5 to 2.5 years old) in semi-structured ways.  The dataset
+plays with toddlers (1.5 to 2.5 years old) in a semi-structured manner.  The dataset
 includes video streams of the faces and the room, audio, physiological data
@@
 of 39 adults recorded over a total of 112 annotated sessions of 3 min in
-average. The participants are instructed to build a tower together with wooden
+average. The participants are instructed to jointly construct a tower using wooden
 blocks. Interestingly, the participants are not allowed to talk to
@@
 following sections, we record much longer interactions (up to 40 minutes) of
-free-play interaction, fostering a wider range of socio-cognitive behaviours.
+free-play interaction, capturing a wider range of socio-cognitive behaviours.
 
-\section{The Plymouth Interacting Social Robots Dataset (PInSoRo)}
+\section{The P[XXXXX] Interacting Social Robots Dataset (PInSoRo)}
 
-The PInSoRo Dataset comprises of natural
+The PInSoRo Dataset is comprised of natural
 social interactions arising between children, recorded with machine learning
@@
     Children are facing each other and sit on cushions. Each child wears a
-    bright sport bib, either purple or yellow, to facilitate later
+    bright sports bib, either purple or yellow, to facilitate later
     identification.}
@@
 
-The acquisition of such a dataset requires first the definition of a suitable
-task for the participants to perform while they are recorded. The nature of
+The acquisition of such a dataset requires a suitable
+task for the participants to perform while they are being recorded. The nature of
 this task influences in fundamental ways the kind of interactions that might be
@@
     \item elicit a large range of interaction situations;
-    \item foster rich multi-modal interaction: simultaneous speech, gesture, and gaze
-        behaviours are to be observed;
+    \item foster rich multi-modal interaction, such as simultaneous speech, gesture, and gaze
+        behaviours;
     \item evidence social dynamics, such as coupling, joint attention, turn-taking.
@@
 The task is designed so that children can engage in open-ended and non-directive play, yet it is
-sufficiently defined to be suitable for recording, and reproducible in comparable conditions.
+sufficiently constrained to be suitable for recording, and allows the reproduction of social behaviour by an artificial agent in comparable conditions.
 
@@
 
-In order to constraint the dataset to a tractable domain, the task is limited to
+In order to constrain the dataset to a tractable domain, the task is limited to
 a dyadic face-to-face interaction.  The resulting interaction space is
-nevertheless high-dimensional, with two main axes: the game-related actions
+nevertheless high-dimensional, with two main foci: the game-related actions
 (drawing, manipulating items, telling stories) and the social interactions
 (monitoring the other's actions, discussing and agreeing on joint actions,
-helping/hostile behaviours, etc.).
+supportive/antagonistic behaviours, etc.).
 
@@
 the children decide themselves when to stop, the duration of each play episode varies ($M=20m51s,
-SD=10m40s$). It is however capped to a maximum of 40 minutes.
+SD=10m40s$). It is however capped at a maximum of 40 minutes.
 
@@
 and equipped with a fast 1TB SSD hard-drive. The computer is held horizontally in
-a custom aluminium frame standing 26cm above the ground
+a custom aluminium frame standing 26cm above the floor
 (visible in Fig.~\ref{fig|freeplay}). All the cameras are plugged directly over
@@
 
-The children faces are recorded from two short range Intel RealSense SR300
+The children's faces are recorded using two short range Intel RealSense SR300
 RGB-D cameras (0.2m to 1.2m) placed at the corners of the the sandtray
@@
 \paragraph{Software}
-The sandbox is implemented with two frameworks: the Qt's QtQuick framework
+The sandbox is implemented using two software frameworks: the Qt's QtQuick framework
 for the graphical interface of the game, and the \emph{Robot Operating System}
@@
 
-The experimental setup is visible in Fig.~\ref{fig|freeplay}, left.  The
+The experimental setup is shown in Fig.~\ref{fig|freeplay}, left.  The
 experimenter stays in the room, visible to the child, during the whole
@@
 
-We recruited normally developing children, from local primary schools and
+We recruited typically developing children, from local primary schools and
 nurseries. The children are aged 4 to 8 ($M=5.8$). So far, 50 children (23
@@
 The following acquisition procedure has been followed with all participants. To
-ease later identification, each child was given a colourful (purple or yellow)
-sport bib to wear.
+ease later identification, each child was given a brightly coloured (purple or yellow)
+sports bib to wear.
 
@@
                                   & \tabitem place children on cushions \\ 
-                                  & \tabitem give them the yellow/purple sport
+                                  & \tabitem give them the yellow/purple sports
                                   bibs and make sure their hair is tied \\ 
@@
 Due to ethical and data protection regulations, the dataset is however made
-available in two forms: a public, Creative-Commons licensed, version that do
+available in two forms: a public, Creative Commons licensed, version that does
 not include any video material of the children (no video streams, audio

### 2017-06-28 paper.tex
Updated github links to freeplay-sandbox organisation

@@
 RGB-D cameras (0.2m to 1.2m) placed at the corners of the the sandtray
-(Figs.~\ref{fig|rviz} and~\ref{fig|setup}) and tilted to face the children. The
+(Fig.~\ref{fig|freeplay}) and tilted to face the children. The
 cameras are rigidly mounted on custom 3D-printed brackets. This enables a
@@
 and the children's touches. The game interface is open-source and available online:
-\url{https://github.com/severin-lemaignan/freeplay-sandbox-qt}.
+\url{https://github.com/freeplay-sandbox/qt-gui}.
 
@@
 The ROS data acquisition pipeline is open-source as well, and available from
-\url{https://github.com/severin-lemaignan/freeplay-sandbox-ros}.
+\url{https://github.com/freeplay-sandbox/core}.
 
@@
 participants. The supervisor is available online as well:
-\url{https://github.com/severin-lemaignan/freeplay-sandbox-supervisor}.
+\url{https://github.com/freeplay-sandbox/web-supervisor}.
 
@@
 dataset\footnote{Available online from
-\url{https://github.com/severin-lemaignan/freeplay-sandbox-analysis}.} to
+\url{https://github.com/freeplay-sandbox/analysis}.} to
 republish registered streams and the corresponding 3D point-clouds (as seen in

### 2017-06-28 paper.tex
Anonymisation

@@
 
-%Indeed, as we have previously argued~\cite{lemaignan2016towards}, 
+%anon Indeed, as we have previously argued~\cite{lemaignan2016towards}, 
 We believe
@@
 
-\section{The P[XXXXX] Interacting Social Robots Dataset (PInSoRo)}
+%anon \section{The Plymouth Interacting Social Robots Dataset (PInSoRo)}
+\section{The [hidden for blind review] Interacting Social Robots Dataset (PInSoRo)}
 
@@
 
-All data has been collected by researchers at the Plymouth University, under a
+All data has been collected by researchers at
+%anon the Plymouth University
+[hidden for blind review], under a
 protocol approved by the university ethics committee. The parents of the

### 2017-06-28 paper.tex
Submitted version

@@
 The children's faces are recorded using two short range Intel RealSense SR300
-RGB-D cameras (0.2m to 1.2m) placed at the corners of the the sandtray
+RGB-D cameras (0.2m to 1.2m) placed at the corners of the sandtray
 (Fig.~\ref{fig|freeplay}) and tilted to face the children. The
@@
 \emph{(1-2 min)}                  & \\ \midrule
-\bf Items placement game          & \emph{(unrelated task on spatial reasoning -- as this task uses the same graphical environment as the main free-play task, it helps the children to familiarise themselves with the game environment)} \\
+\bf Items placement game          & \emph{(unrelated task on spatial reasoning -- as this task uses the same graphical environment as the main free-play task, it helps however the children to familiarise themselves with the game environment)} \\
 \emph{(5 to 10 min)}              &  \\ \midrule
@@
 
-The last part of the dataset consist in annotations of the children's social
+The last part of the dataset consists in annotations of the children's social
 behaviours. These behaviours are either automatically computed and coded
 (mutual gaze, for instance), or are human-coded. The annotations are all
-timestamped and integrated as an independent datastream in the dataset.
+timestamped and integrated as an supplementary datastream in the dataset.
 
@@
 System} (SCCS) proposed by Olswang~\etal\cite{olswang2006reliability}.
-These annotations are performed post-hoc, with a dedicated tool, able to replay synchronously
-the main datastreams (Fig.~\ref{fig|analysis}).
-
 The SCCS consists in 6 mutually exclusive constructs characterising social
@@
 classification output that a network could be trained to recognise.
+
+The SCCS coding is performed post-hoc with the help of a dedicated tool, able
+to replay synchronously the main datastreams (Fig.~\ref{fig|analysis}) while
+the human coder annotates the interaction.
 
@@
 
-\begin{figure}[b!]
+\begin{figure}[th!]
     \centering
@@
     is played back synchronously (faces and audio of the two children,
-    interaction occurring on the interactive table, etc.) while the human coder
-    annotate the record timeline with social situations of interest.}
+    interaction occurring on the interactive table, environment camera) while the human coder
+    annotates the record timeline with social situations of interest.}
     \label{fig|analysis}
@@
 available online: \url{https://freeplay-sandbox.github.io/}. A link to the
-complete dataset will be made available in the final version of the article.}.
+complete dataset will be made available with the final version of the article.}.
 Due to ethical and data protection regulations, the dataset is however made
@@
 During the acquisition campaign for the PInSoRo dataset, we have also recorded a
-\emph{child-robot} interaction baseline. We were using the exact same protocol
+\emph{child-robot} interaction baseline. We use the exact same protocol
 as described above, only with one child replaced by a Nao robot.
 
-The robot was fully autonomous, and programmed to exhibit a non-social behaviour
-(its action policy was simply to repeatedly move interactive items to
-pre-defined zones, entirely ignoring its child partner).
+The robot is fully autonomous, and programmed to exhibit a \emph{non-social}
+behaviour (its action policy is simply to repeatedly move interactive items to
+pre-defined zones, entirely ignoring its child partner).  This behaviour (and
+the resulting reactions of the children to a robot acting in a non-social way)
+represents an experimental baseline against which future work on social
+human-robot interaction can be compared.
 
-As such, this behaviour (and the resulting reaction of the children to a
-robot acting in a non-social way) represent an experimental baseline against which
-future work can be compared.
+This second dataset is public as well, and is made available to the interested
+researchers under the same conditions as the child-child interaction dataset.
 
@@
 
-We introduce a novel dataset for the study of social cognition. The dataset
-consists in dyadic free-play episodes, with children aged 4 to 8.
+In this article, we have introduced a novel dataset for the study of early
+social cognition. The dataset consists in dyadic free-play episodes with
+children aged 4 to 8. The interactions are scaffolded by a novel, playful task,
+and a large range of social behaviours have been observed: pro-social behaviours
+(for instance, helping behaviours, negotiations, natural turn-taking),
+antagonistic behaviours (for instance, disrupting behaviours, refusal to
+cooperate), assertive behaviours (dominant behaviours, manipulative behaviours)
+and numerous episodes of off-topic actions and interactions.
 
-The dataset is built with machine learning applications in mind, and
-particularly, applications to socio-cognitive robotics.
+For a robot to recognise such behaviours and situations would be of great
+interest to the human-robot interaction community as it would effectively
+increase the robot's situation awareness. This would in turn enable the robot to
+respond appropriately to the complex social situations encountered in real-world
+scenarios.
 
-Finally, the social setting (two players facing each other) implicitly invites
-social interaction: onlooker play, for instance, is unlikely to be observed as
-the children are physically engaged with the game and the other participant
-(they are sitting in front of game, close to each other). The three higher forms
-of social play listed by Parten (parallel play, associative play and cooperative
-play) are however possible with this setup.
+Our dataset is open, and built with machine learning applications in mind, and
+particularly, applications to socio-cognitive robotics: we hope it will
+effectively contribute to bridge the gap between machine learning and social
+robotics, and support the research efforts of our community.
+
+%Finally, the social setting (two players facing each other) implicitly invites
+%social interaction: onlooker play, for instance, is unlikely to be observed as
+%the children are physically engaged with the game and the other participant
+%(they are sitting in front of game, close to each other). The three higher forms
+%of social play listed by Parten (parallel play, associative play and cooperative
+%play) are however possible with this setup.
 
